---
title: "CF Interactions"
author: "Reed Stubbendieck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(circlize)
library(ComplexHeatmap)
library(ComplexUpset)
library(cowplot)
library(ggtext)
library(lme4)
library(lmerTest)
library(multcompView)
library(parallel)
library(parallelDist)
library(patchwork)
library(pvclust)
library(readxl)
library(scales)
library(tidyverse)
library(vegan)

set.seed(8675309)
```

## Load Data

```{r load_data, include=FALSE}
cf_microbiome_metadata <- read.csv("./rawData/16S_rRNA_gene_amplicon_studies/16S_CF_microbiome.csv")
cf_isolates <- read_csv("./rawData/isolate_identification/isolate_identification.csv")
bioassay_data <- read_csv("./rawData/cross-streak/cross-streak_inhibition_data.csv")

cf_pathogens <- c("Pseudomonas", "Staphylococcus", "Burkholderia",
                  "Stenotrophomonas", "Achromobacter", "Haemophilus",
                  "Escherichia")

files <- list.files("derivedData/mothur_output", pattern = "\\.tax\\.summary$", full.names = TRUE)

lcms_dir  <- file.path("rawData", "lcms", "overall")
lcms_diff <- read_excel(file.path(lcms_dir, "XCMS-diffreport-MultiClass.xlsx"), sheet = 1)
lcms_meta <- read_excel(file.path("rawData", "lcms", "pseudomonas_isolate_metadata.xlsx"), sheet = 1)
names(lcms_meta)[names(lcms_meta) == "Inhibition_Z-Score"] <- "inhibition_z"
if (!"NB1386" %in% lcms_meta$isolate)                       # NB1386 absent from the sheet
  lcms_meta <- bind_rows(lcms_meta, data.frame(isolate = "NB1386", host = "HID08", inhibition_z = -0.86))
lcms_meta$inhibition <- ifelse(lcms_meta$inhibition_z >= 0, "Yes", "No")

# --- LC-MS/MS total ion chromatograms (cloud plot) ---
lcms_env  <- new.env(); load(file.path(lcms_dir, "xcms3_TIC.Rda"), envir = lcms_env)
lcms_tics <- lcms_env$TICS.rtcor
```

## 16S rRNA gene amplicon sequencing processing

```{r CF_microbiota_processing, include=FALSE}
parse_tax_summary <- function(filepath) {
  read_tsv(filepath, col_types = cols(.default = "c"), show_col_types = FALSE) %>%
    mutate(taxlevel = as.integer(taxlevel), reads = as.integer(total)) %>%
    filter(taxlevel == 6, !str_detect(taxon, "unclassified")) %>%
    dplyr::select(taxon, reads) %>%
    mutate(sample = sub("\\.tax\\.summary$", "", basename(filepath)))
}

get_total_reads <- function(filepath) {
  read_tsv(filepath, col_types = cols(.default = "c"), show_col_types = FALSE, n_max = 2) %>%
    filter(as.integer(taxlevel) == 0) %>%
    transmute(sample = sub("\\.tax\\.summary$", "", basename(filepath)),
              total_reads = as.integer(total))
}

keep_samples <- map_dfr(files, get_total_reads) %>%
  filter(total_reads >= 1000) %>%
  pull(sample)

all_genera <- map_dfr(files, parse_tax_summary) %>%
  filter(sample %in% keep_samples)

metacommunity <- all_genera %>%
  group_by(genus = taxon) %>%
  summarize(total_reads = sum(reads), .groups = "drop") %>%
  mutate(rel_abund = round(total_reads / sum(total_reads)*100),2) %>%
  arrange(desc(rel_abund)) %>%
  mutate(genus = if_else(row_number() <= 20, genus, "Other")) %>%
  group_by(genus) %>%
  summarize(rel_abund = sum(rel_abund), .groups = "drop") %>%
  arrange(desc(rel_abund))

genus_matrix <- all_genera %>%
  pivot_wider(id_cols = sample, names_from = taxon,
              values_from = reads, values_fill = 0) %>%
  column_to_rownames("sample") %>%
  as.matrix()

row_totals <- rowSums(genus_matrix)
row_totals[row_totals == 0] <- NA
genus_matrix[sweep(genus_matrix, 1, row_totals, FUN = "/") < 0.001] <- 0

genus_matrix <- genus_matrix[,
  names(which(colSums(genus_matrix > 0) >= ceiling(0.01 * nrow(genus_matrix))))
]

table1_genera <- c("Streptococcus", "Pseudomonas", "Staphylococcus", "Prevotella",
                   "Haemophilus", "Veillonella", "Rothia", "Stenotrophomonas",
                   "Neisseria", "Granulicatella", "Achromobacter", "Gemella",
                   "Porphyromonas", "Burkholderia", "Fusobacterium", "Actinomyces",
                   "Schaalia", "Alloprevotella", "Escherichia")

overall_rel_abund <- colSums(genus_matrix) / sum(genus_matrix) * 100
prevalence_pct    <- colSums(genus_matrix > 0) / nrow(genus_matrix) * 100

stopifnot(all(table1_genera %in% colnames(genus_matrix)))
keep_genera <- table1_genera

genus_matrix_filt <- genus_matrix[, keep_genera, drop = FALSE]

# report the Table 1 quantities for the genera used
print(data.frame(genus = keep_genera,
                 overall_rel_abund = round(overall_rel_abund[keep_genera], 1),
                 prevalence_pct    = round(prevalence_pct[keep_genera], 1),
                 row.names = NULL))

write.csv(genus_matrix_filt, "derivedData/genus_matrix_filt.csv")
```

```{r metacommunity_table, include=FALSE}
n_samples   <- length(keep_samples)
grand_total <- sum(all_genera$reads)

per_sample_ra <- all_genera %>%
  group_by(sample) %>%
  mutate(sample_total = sum(reads)) %>%
  ungroup() %>%
  mutate(ra = reads / sample_total * 100)

metacommunity_table <- per_sample_ra %>%
  group_by(genus = taxon) %>%
  summarize(
    metacommunity_ra = sum(reads) / grand_total * 100,
    n_prev_01        = sum(ra >= 0.1),
    n_prev_1         = sum(ra >= 1),
    n_dominant       = sum(ra > 50),
    mean_ra          = mean(ra[ra > 0]),
    median_ra        = median(ra[ra > 0]),
    q25_ra           = quantile(ra[ra > 0], 0.25),
    q75_ra           = quantile(ra[ra > 0], 0.75),
    mean_ra_dom      = ifelse(sum(ra > 50) == 0, NA_real_,
                              mean(ra[ra > 50])),
    median_ra_dom    = ifelse(sum(ra > 50) == 0, NA_real_,
                              median(ra[ra > 50])),
    q25_ra_dom       = ifelse(sum(ra > 50) == 0, NA_real_,
                              quantile(ra[ra > 50], 0.25)),
    q75_ra_dom       = ifelse(sum(ra > 50) == 0, NA_real_,
                              quantile(ra[ra > 50], 0.75)),
    .groups = "drop"
  ) %>%
  arrange(desc(metacommunity_ra)) %>%
  filter(metacommunity_ra >= 0.5) %>%
  mutate(
    metacommunity_ra  = round(metacommunity_ra, 1),
    mean_ra           = round(mean_ra, 1),
    median_ra         = round(median_ra, 1),
    range_ra          = paste0(round(q25_ra, 1), "-", round(q75_ra, 1)),
    prev_01_pct       = round(n_prev_01 / n_samples * 100, 1),
    prev_1_pct        = round(n_prev_1  / n_samples * 100, 1),
    dominant_pct      = round(n_dominant / n_samples * 100, 1),
    conditional_dominance  = round(
      ifelse(n_prev_1 == 0, NA_real_, n_dominant / n_prev_1 * 100), 1),
    mean_ra_dom       = round(mean_ra_dom, 1),
    median_ra_dom     = round(median_ra_dom, 1),
    range_ra_dom      = ifelse(is.na(q25_ra_dom), NA_character_,
                               paste0(round(q25_ra_dom, 1), "-",
                                      round(q75_ra_dom, 1)))
  ) %>%
  dplyr::select(
    genus, metacommunity_ra,
    n_prev_01, prev_01_pct,
    n_prev_1,  prev_1_pct,
    n_dominant, dominant_pct, conditional_dominance,
    mean_ra, median_ra, range_ra,
    mean_ra_dom, median_ra_dom, range_ra_dom
  )

print(metacommunity_table, n = 30)

metacommunity_csv <- metacommunity_table %>%
  mutate(
    median_iqr     = paste0(median_ra, " [", range_ra, "]"),
    median_iqr_dom = ifelse(is.na(median_ra_dom), NA_character_,
                            paste0(median_ra_dom, " [", range_ra_dom, "]"))
  ) %>%
  dplyr::select(
    Genus                              = genus,
    `Metacommunity RA (%)`             = metacommunity_ra,
    `Mean RA (%)`                      = mean_ra,
    `Median RA % [IQR]`                = median_iqr,
    `Prevalence >=1% (%)`              = prev_1_pct,
    `Dominant (>50%) (%)`              = dominant_pct,
    `Dominant given Prev >=1% (%)`     = conditional_dominance,
    `Mean RA when Dominant (%)`        = mean_ra_dom,
    `Median RA when Dominant % [IQR]`  = median_iqr_dom
  )

write.csv(metacommunity_csv, "derivedData/table_metacommunity.csv", row.names = FALSE)
```

## Conditional dominance tests: pathogen vs oropharyngeal

```{r dominance_tests, include=FALSE}
metacommunity_genera <- metacommunity_table$genus

per_sample_ra_lab <- per_sample_ra %>%
  filter(taxon %in% metacommunity_genera) %>%
  mutate(group = ifelse(taxon %in% cf_pathogens, "pathogen", "oropharyngeal"))

conditional_dominance_per_genus <- per_sample_ra_lab %>%
  group_by(taxon, group) %>%
  summarize(n_prev_1   = sum(ra >= 1),
            n_dominant = sum(ra > 50),
            .groups    = "drop") %>%
  filter(n_prev_1 > 0) %>%
  mutate(conditional_dominance = n_dominant / n_prev_1 * 100)

q1_wilcox <- wilcox.test(conditional_dominance ~ group, data = conditional_dominance_per_genus,
                         exact = FALSE)
print(q1_wilcox)
conditional_dominance_per_genus %>%
  group_by(group) %>%
  summarize(n_genera    = n(),
            median_rate = median(conditional_dominance),
            q25         = quantile(conditional_dominance, 0.25),
            q75         = quantile(conditional_dominance, 0.75)) %>%
  print()

dominant_obs <- per_sample_ra_lab %>% filter(ra > 50)

q2_wilcox <- wilcox.test(ra ~ group, data = dominant_obs, exact = FALSE)
print(q2_wilcox)
dominant_obs %>%
  group_by(group) %>%
  summarize(n_dominant_samples = n(),
            median_ra          = median(ra),
            q25                = quantile(ra, 0.25),
            q75                = quantile(ra, 0.75)) %>%
  print()

pathogen_dominant <- dominant_obs %>%
  filter(group == "pathogen") %>%
  mutate(taxon = factor(taxon))

q3_kw <- kruskal.test(ra ~ taxon, data = pathogen_dominant)
print(q3_kw)
pathogen_dominant %>%
  group_by(taxon) %>%
  summarize(n_dominant_samples = n(),
            median_ra          = median(ra),
            q25                = quantile(ra, 0.25),
            q75                = quantile(ra, 0.75)) %>%
  arrange(desc(median_ra)) %>%
  print()
```

## NMDS

```{r nmds, include=FALSE}
rel_abund_matrix <- sweep(genus_matrix, 1, rowSums(genus_matrix), "/")
rel_abund_matrix[is.nan(rel_abund_matrix)] <- 0

dir.create("r_objects", showWarnings = FALSE)

if (file.exists("r_objects/bray_dist.rds") && file.exists("r_objects/nmds.rds")) {
  bray_dist <- readRDS("r_objects/bray_dist.rds")
  nmds      <- readRDS("r_objects/nmds.rds")
} else {
  bray_dist <- parDist(rel_abund_matrix, method = "bray")

  n_cores <- max(1, parallel::detectCores() - 1)
  cl      <- makeCluster(n_cores)
  clusterEvalQ(cl, library(vegan))
  clusterExport(cl, "bray_dist")

  starts <- parLapply(cl, seq_len(20), function(i) monoMDS(bray_dist, k = 2, maxit = 500))
  stopCluster(cl)

  nmds <- starts[[which.min(sapply(starts, `[[`, "stress"))]]

  saveRDS(bray_dist, "r_objects/bray_dist.rds")
  saveRDS(nmds,      "r_objects/nmds.rds")
}

nmds_scores <- as.data.frame(scores(nmds, display = "sites")) %>%
  rownames_to_column("Run") %>%
  left_join(cf_microbiome_metadata, by = "Run")

nmds_stress <- nmds$stress
```

### Beta Diversity: PERMANOVA

```{r permanova, include=FALSE}
if (file.exists("r_objects/permanova_combined.rds")) {
  permanova_combined <- readRDS("r_objects/permanova_combined.rds")
} else {
  set.seed(8675309)
  N_PERM <- 9999

  bray_mat <- as.matrix(bray_dist)

  cf_microbiome_metadata$Age_bin <- cut(cf_microbiome_metadata$Age,
    breaks = c(-Inf, 2, 12, 18, 40, Inf), right = FALSE,
    labels = c("Infant", "Child", "Adolescent", "Adult", "Senior"))

  cf_microbiome_metadata$FEV1_bin <- cut(cf_microbiome_metadata$FEV1,
    breaks = c(-Inf, 30, 50, 80, Inf), right = FALSE,
    labels = c("Very Severe", "Severe", "Moderate", "Mild"))

  keep_idx <- function(meta, variable) {
    x <- meta[[variable]]
    which(!is.na(x) & !(is.character(x) & x == ""))
  }

  tech_vars       <- c("BioProject", "Region")
  bio_vars        <- c("Sex", "Age", "FEV1", "Sample_Type")
  binned_vars     <- c("Age_bin", "FEV1_bin")
  partial_vars    <- c(bio_vars, binned_vars)
  partial_st_vars <- c("Sex", "Age_bin", "FEV1_bin")

  tasks <- c(
    lapply(c(tech_vars, bio_vars, binned_vars),
           function(v) list(type = "unadj",           variable = v)),
    lapply(partial_vars,
           function(v) list(type = "partial_br",      variable = v)),
    lapply(partial_st_vars,
           function(v) list(type = "partial_brst",    variable = v)),
    list(list(type = "joint_technical",
              variable = "BioProject + Sample_Type"))
  )

  n_cores <- max(1, parallel::detectCores() - 1)
  cl      <- parallel::makeCluster(n_cores)
  parallel::clusterEvalQ(cl, library(vegan))
  parallel::clusterExport(cl, c("cf_microbiome_metadata", "bray_mat",
                                 "keep_idx", "N_PERM"))

  results <- parallel::parLapply(cl, tasks, function(task) {
    meta <- cf_microbiome_metadata
    v    <- task$variable

    if (task$type == "unadj") {
      idx      <- keep_idx(meta, v)
      dist_sub <- as.dist(bray_mat[idx, idx])
      meta_sub <- meta[idx, , drop = FALSE]
      f <- as.formula(paste("dist_sub ~", v))
      environment(f) <- environment()
      res <- as.data.frame(adonis2(f, data = meta_sub, permutations = N_PERM))
      res <- res[!rownames(res) %in% c("Residual", "Total"), , drop = FALSE]
      return(data.frame(variable = v, n = length(idx), Df = res$Df,
                        R2 = round(res$R2, 4), F = round(res$F, 3),
                        p.value = res[["Pr(>F)"]], model = "unadjusted",
                        row.names = NULL, check.names = FALSE))
    }

    if (task$type == "partial_br") {
      idx      <- keep_idx(meta, v)
      dist_sub <- as.dist(bray_mat[idx, idx])
      meta_sub <- meta[idx, , drop = FALSE]
      f <- as.formula(paste("dist_sub ~ BioProject +", v))
      environment(f) <- environment()
      res <- as.data.frame(adonis2(f, data = meta_sub,
                                   permutations = N_PERM, by = "margin"))
      res <- res[v, , drop = FALSE]
      return(data.frame(variable = v, n = length(idx), Df = res$Df,
                        R2 = round(res$R2, 4), F = round(res$F, 3),
                        p.value = res[["Pr(>F)"]],
                        model = "partial (| BioProject)",
                        row.names = NULL, check.names = FALSE))
    }

    if (task$type == "partial_brst") {
      idx      <- keep_idx(meta, v)
      idx      <- idx[!is.na(meta$Sample_Type[idx]) & meta$Sample_Type[idx] != ""]
      dist_sub <- as.dist(bray_mat[idx, idx])
      meta_sub <- meta[idx, , drop = FALSE]
      f <- as.formula(paste("dist_sub ~ BioProject + Sample_Type +", v))
      environment(f) <- environment()
      res <- as.data.frame(adonis2(f, data = meta_sub,
                                   permutations = N_PERM, by = "margin"))
      res <- res[v, , drop = FALSE]
      return(data.frame(variable = v, n = length(idx), Df = res$Df,
                        R2 = round(res$R2, 4), F = round(res$F, 3),
                        p.value = res[["Pr(>F)"]],
                        model = "partial (| BioProject + Sample_Type)",
                        row.names = NULL, check.names = FALSE))
    }

    if (task$type == "joint_technical") {
      idx      <- which(!is.na(meta$Sample_Type) & meta$Sample_Type != "")
      dist_sub <- as.dist(bray_mat[idx, idx])
      meta_sub <- meta[idx, , drop = FALSE]
      res <- as.data.frame(adonis2(
        dist_sub ~ BioProject + Sample_Type,
        data = meta_sub, permutations = N_PERM))
      res <- res["Model", , drop = FALSE]
      return(data.frame(variable = v, n = length(idx), Df = res$Df,
                        R2 = round(res$R2, 4), F = round(res$F, 3),
                        p.value = res[["Pr(>F)"]], model = "joint technical",
                        row.names = NULL, check.names = FALSE))
    }
  })

  parallel::stopCluster(cl)

  permanova_combined       <- do.call(rbind, results)
  permanova_combined$p.adj <- p.adjust(permanova_combined$p.value, method = "BH")

  saveRDS(permanova_combined, "r_objects/permanova_combined.rds")
}

print(permanova_combined, digits = 3, row.names = FALSE)

technical_joint_R2 <- permanova_combined$R2[permanova_combined$model == "joint technical"]

if (file.exists("r_objects/permanova_subsampling_summary.rds")) {
  permanova_subsampling_summary <- readRDS("r_objects/permanova_subsampling_summary.rds")
} else {
  set.seed(8675309)
  N_PERM_SUB <- 999
  N_REPS     <- 1000
  SUB_N      <- 100
  bray_mat <- as.matrix(bray_dist)

  if (!"Age_bin" %in% colnames(cf_microbiome_metadata)) {
    cf_microbiome_metadata$Age_bin <- cut(cf_microbiome_metadata$Age,
      breaks = c(-Inf, 2, 12, 18, 40, Inf), right = FALSE,
      labels = c("Infant", "Child", "Adolescent", "Adult", "Senior"))
  }
  if (!"FEV1_bin" %in% colnames(cf_microbiome_metadata)) {
    cf_microbiome_metadata$FEV1_bin <- cut(cf_microbiome_metadata$FEV1,
      breaks = c(-Inf, 30, 50, 80, Inf), right = FALSE,
      labels = c("Very Severe", "Severe", "Moderate", "Mild"))
  }

  keep_idx_fn <- function(meta, variable) {
    x <- meta[[variable]]
    which(!is.na(x) & !(is.character(x) & x == ""))
  }

  sub_vars <- c("Sex", "Age_bin", "FEV1_bin")
  sub_tasks <- expand.grid(variable = sub_vars, rep = seq_len(N_REPS),
                           stringsAsFactors = FALSE) %>%
    apply(1, function(row) list(variable = row[["variable"]],
                                rep = as.integer(row[["rep"]])))

  n_cores <- max(1, parallel::detectCores() - 1)
  cl <- parallel::makeCluster(n_cores)
  parallel::clusterEvalQ(cl, library(vegan))
  parallel::clusterExport(cl, c("cf_microbiome_metadata", "bray_mat",
                                "keep_idx_fn", "N_PERM_SUB", "SUB_N"),
                          envir = environment())

  sub_results <- parallel::parLapply(cl, sub_tasks, function(task) {
    meta <- cf_microbiome_metadata
    v    <- task$variable
    idx_full <- keep_idx_fn(meta, v)
    if (length(idx_full) < SUB_N) {
      return(list(variable = v, rep = task$rep, R2 = NA, pval = NA))
    }
    set.seed(task$rep * 1000 + SUB_N +
             as.integer(factor(v, levels = c("Sex","Age_bin","FEV1_bin"))))
    idx <- sample(idx_full, SUB_N)
    dist_sub <- as.dist(bray_mat[idx, idx])
    meta_sub <- meta[idx, ]
    f <- as.formula(paste("dist_sub ~", v))
    environment(f) <- environment()
    res <- tryCatch(adonis2(f, data = meta_sub, permutations = N_PERM_SUB),
                    error = function(e) NULL)
    if (is.null(res)) return(list(variable = v, rep = task$rep, R2 = NA, pval = NA))
    res <- as.data.frame(res)
    list(variable = v, rep = task$rep,
         R2 = res["Model", "R2"], pval = res["Model", "Pr(>F)"])
  })
  parallel::stopCluster(cl)

  permanova_subsampling <- do.call(rbind, lapply(sub_results, function(r)
    data.frame(variable = r$variable, N = SUB_N, rep = r$rep,
               R2 = r$R2, pval = r$pval,
               sig = !is.na(r$pval) & r$pval < 0.05,
               stringsAsFactors = FALSE)))
  saveRDS(permanova_subsampling, "r_objects/permanova_subsampling.rds")

  permanova_subsampling_summary <- permanova_subsampling %>%
    group_by(variable, N) %>%
    summarize(frac_sig = mean(sig, na.rm = TRUE),
              mean_R2  = mean(R2,  na.rm = TRUE),
              sd_R2    = sd(R2,   na.rm = TRUE),
              .groups  = "drop")
  saveRDS(permanova_subsampling_summary, "r_objects/permanova_subsampling_summary.rds")
}

print(as.data.frame(permanova_subsampling_summary))
```

## Alpha Diversity vs FEV1

```{r alpha_diversity_fev1, include=FALSE}
parse_tax_genus <- function(filepath) {
  read_tsv(filepath, col_types = cols(.default = "c"), show_col_types = FALSE) %>%
    mutate(taxlevel = as.integer(taxlevel), reads = as.integer(total)) %>%
    filter(taxlevel == 6, !str_detect(taxon, "unclassified")) %>%
    dplyr::select(taxon, reads) %>%
    mutate(sample = sub("\\.tax\\.summary$", "", basename(filepath)))
}

sample_totals <- map_dfr(files, function(f) {
  read_tsv(f, col_types = cols(.default = "c"), show_col_types = FALSE, n_max = 2) %>%
    filter(as.integer(taxlevel) == 0) %>%
    transmute(sample = sub("\\.tax\\.summary$", "", basename(f)),
              total_reads = as.integer(total))
}) %>% filter(total_reads >= 1000)

good_samples <- sample_totals$sample

mat_genus <- map_dfr(files, parse_tax_genus) %>%
  filter(sample %in% good_samples) %>%
  pivot_wider(id_cols = sample, names_from = taxon,
              values_from = reads, values_fill = 0) %>%
  column_to_rownames("sample") %>%
  as.matrix()

alpha_df <- data.frame(
  sample  = rownames(mat_genus),
  shannon = vegan::diversity(mat_genus, index = "shannon"),
  stringsAsFactors = FALSE
)

alpha_df <- alpha_df %>%
  left_join(cf_microbiome_metadata %>% dplyr::select(Run, FEV1, Age, BioProject, Sample_Type, Region),
            by = c("sample" = "Run")) %>%
  filter(!is.na(FEV1), FEV1 != "") %>%
  mutate(FEV1 = as.numeric(FEV1),
         Age  = as.numeric(Age))

alpha_df <- alpha_df %>%
  mutate(FEV1_bin = cut(FEV1,
    breaks = c(-Inf, 30, 50, 80, Inf), right = FALSE,
    labels = c("Very Severe", "Severe", "Moderate", "Mild")))

# Adjust Shannon for Age + BioProject + Sample_Type + Region (residuals + grand mean)
complete <- !is.na(alpha_df$Age) & !is.na(alpha_df$BioProject) &
            !is.na(alpha_df$Sample_Type) & alpha_df$Sample_Type != "" &
            !is.na(alpha_df$Region) & alpha_df$Region != ""
alpha_df$shannon_adj <- NA_real_
fit_s <- lm(shannon ~ Age + BioProject + Sample_Type + Region, data = alpha_df[complete, ])
alpha_df$shannon_adj[complete] <- residuals(fit_s) + mean(alpha_df$shannon[complete])

alpha_adj <- alpha_df %>% filter(!is.na(shannon_adj))

kw_result <- kruskal.test(shannon_adj ~ FEV1_bin, data = alpha_adj)
print(kw_result)

pw_result <- pairwise.wilcox.test(alpha_adj$shannon_adj, alpha_adj$FEV1_bin,
                                   p.adjust.method = "BH", exact = FALSE)
print(pw_result$p.value)
```

## Berger-Parker Dominance

```{r berger_parker_prep, include=FALSE}
pathogen_cols   <- intersect(cf_pathogens, colnames(mat_genus))
pathogen_reads  <- rowSums(mat_genus[, pathogen_cols, drop = FALSE])
bp_dom_vec      <- apply(mat_genus, 1, max) / rowSums(mat_genus)
pathogen_frac   <- pathogen_reads / rowSums(mat_genus)
dominant_genus  <- colnames(mat_genus)[max.col(mat_genus, ties.method = "first")]

bp_df <- data.frame(
  sample           = rownames(mat_genus),
  bp_dom           = bp_dom_vec,
  pathogen_frac    = pathogen_frac,
  dominant_genus   = dominant_genus,
  stringsAsFactors = FALSE
) %>%
  mutate(
    dominant_pathogen = if_else(dominant_genus %in% cf_pathogens,
                                dominant_genus, "Other")
  ) %>%
  left_join(
    cf_microbiome_metadata %>%
      dplyr::select(Run, FEV1, Age, BioProject, Sample_Type, Region),
    by = c("sample" = "Run")
  ) %>%
  filter(!is.na(FEV1), FEV1 != "") %>%
  mutate(
    FEV1     = as.numeric(FEV1),
    Age      = as.numeric(Age),
    FEV1_bin = cut(FEV1,
                   breaks = c(-Inf, 30, 50, 80, Inf), right = FALSE,
                   labels = c("Very Severe", "Severe", "Moderate", "Mild"))
  )

bp_complete <- !is.na(bp_df$Age) &
               !is.na(bp_df$BioProject) &
               !is.na(bp_df$Sample_Type) & bp_df$Sample_Type != "" &
               !is.na(bp_df$Region)      & bp_df$Region      != ""

fit_bp <- lm(bp_dom ~ Age + BioProject + Sample_Type + Region,
             data = bp_df[bp_complete, ])
bp_df$bp_adj <- NA_real_
bp_df$bp_adj[bp_complete] <- residuals(fit_bp) + mean(bp_df$bp_dom[bp_complete])

fit_fev1 <- lm(FEV1 ~ Age + BioProject + Sample_Type + Region,
               data = bp_df[bp_complete, ])
bp_df$FEV1_adj <- NA_real_
bp_df$FEV1_adj[bp_complete] <- residuals(fit_fev1) + mean(bp_df$FEV1[bp_complete])

bp_adj_df <- bp_df %>% filter(!is.na(bp_adj))

n_boot_bp <- 100
n_bp      <- nrow(bp_adj_df)
boot_lines_bp <- do.call(rbind, lapply(seq_len(n_boot_bp), function(b) {
  bi  <- sample(seq_len(n_bp), replace = TRUE)
  fit <- lm(bp_adj_df$FEV1_adj[bi] ~ bp_adj_df$bp_dom[bi])
  data.frame(x    = range(bp_adj_df$bp_dom),
             y    = coef(fit)[1] + coef(fit)[2] * range(bp_adj_df$bp_dom),
             boot = b)
}))

bp_full_lm  <- lm(FEV1 ~ bp_dom + Age + BioProject + Sample_Type + Region,
                  data = bp_df[bp_complete, ])
bp_full_s   <- summary(bp_full_lm)
bp_coef_row <- bp_full_s$coefficients["bp_dom", ]
bp_p_fmt    <- format(bp_coef_row["Pr(>|t|)"], digits = 2, scientific = TRUE)
bp_stat_label <- paste0("italic(P)[adj] == ", bp_p_fmt)

print(bp_full_s$coefficients["bp_dom", ])
sprintf("Adj. R^2 = %.3f", bp_full_s$adj.r.squared)
```

## Interaction Analyses

```{r pairwise_cooccurrence, include=FALSE}
genera   <- colnames(genus_matrix_filt)
n_genera <- length(genera)

log_abund <- log10(genus_matrix_filt + 1)

pairs   <- combn(genera, 2)
n_pairs <- ncol(pairs)
bonf_thresh <- 0.05 / n_pairs

pw_res <- vector("list", n_pairs)
for (k in seq_len(n_pairs)) {
  g1 <- pairs[1, k]; g2 <- pairs[2, k]
  fit <- lm(log_abund[, g1] ~ log_abund[, g2])
  s   <- summary(fit)
  pw_res[[k]] <- data.frame(genus1 = g1, genus2 = g2,
                             bx = coef(fit)[2], pval = s$coefficients[2, 4],
                             stringsAsFactors = FALSE)
}
pairwise_results <- do.call(rbind, pw_res)
pairwise_results$significant <- pairwise_results$pval < bonf_thresh

edge_list <- pairwise_results %>%
  filter(significant) %>%
  mutate(direction = ifelse(bx > 0, "positive", "negative"))

sig_pairs <- pairwise_results %>%
  filter(significant) %>%
  mutate(
    g1_path   = genus1 %in% cf_pathogens,
    g2_path   = genus2 %in% cf_pathogens,
    pair_type = case_when(
      g1_path & g2_path   ~ "pathogen-pathogen",
      g1_path | g2_path   ~ "pathogen-anaerobe",
      TRUE                ~ "anaerobe-anaerobe"
    ),
    direction = ifelse(bx > 0, "positive", "negative")
  )

pair_dir_summary <- sig_pairs %>%
  count(pair_type, direction) %>%
  group_by(pair_type) %>%
  mutate(total = sum(n), pct = 100 * n / total) %>%
  ungroup()

print(as.data.frame(pair_dir_summary))
print(sig_pairs %>% count(pair_type))
```

```{r pathogenic_interaction_model, include=FALSE}
fev1_meta <- cf_microbiome_metadata %>%
  filter(Run %in% rownames(genus_matrix_filt), !is.na(FEV1), FEV1 != "")
fev1_runs <- fev1_meta$Run
fev1_vals <- as.numeric(fev1_meta$FEV1)
log_fev1  <- log_abund[fev1_runs, , drop = FALSE]

int_res <- vector("list", n_pairs)
for (k in seq_len(n_pairs)) {
  g1 <- pairs[1, k]; g2 <- pairs[2, k]
  pn <- log_fev1[, g1]; px <- log_fev1[, g2]
  fit <- lm(fev1_vals ~ pn + px + pn:px)
  s   <- summary(fit)
  coefs <- s$coefficients
  bi   <- if ("pn:px" %in% rownames(coefs)) coefs["pn:px", "Estimate"] else NA
  pval <- if ("pn:px" %in% rownames(coefs)) coefs["pn:px", "Pr(>|t|)"] else NA
  int_res[[k]] <- data.frame(genus1 = g1, genus2 = g2, bi = bi,
                              pval_interaction = pval, stringsAsFactors = FALSE)
}
interaction_df <- do.call(rbind, int_res) %>% filter(!is.na(bi))
interaction_df$fdr_interaction <- p.adjust(interaction_df$pval_interaction, method = "BH")
interaction_df$sig_interaction <- interaction_df$fdr_interaction < 0.05

covar_pairs <- pairwise_results[pairwise_results$significant, c("genus1", "genus2")]
sig_taxa    <- unique(c(covar_pairs$genus1, covar_pairs$genus2))

PRES_THRESH   <- 1     # a genus counts as present at >=1% within-sample relative abundance
pres_ra       <- 100 * genus_matrix_filt / rowSums(genus_matrix_filt)
pres_ra[!is.finite(pres_ra)] <- 0
pres_matrix   <- pres_ra[fev1_runs, , drop = FALSE] >= PRES_THRESH
n_interacting <- rowSums(pres_matrix[, covar_pairs$genus1, drop = FALSE] &
                         pres_matrix[, covar_pairs$genus2, drop = FALSE])
bioproject    <- fev1_meta$BioProject
sample_type   <- fev1_meta$Sample_Type
region        <- fev1_meta$Region

fev1_age_raw <- fev1_meta %>%
  mutate(Age_bin = case_when(
    is.na(Age)           ~ "NA",
    Age < 12             ~ "0\u201311",
    Age >= 12 & Age < 18 ~ "12\u201317",
    Age >= 18 & Age < 40 ~ "18\u201339",
    Age >= 40            ~ "40+"
  )) %>% pull(Age_bin)

age_levels_raw <- c("0\u201311", "12\u201317", "18\u201339", "40+", "NA")
age_counts     <- table(factor(fev1_age_raw, levels = age_levels_raw))
age_label_map  <- setNames(paste0(names(age_counts), " (N=", age_counts, ")"),
                           names(age_counts))
fev1_age <- age_label_map[fev1_age_raw]

fev1_age_vals <- as.numeric(fev1_meta$Age)
complete_idx  <- !is.na(fev1_age_vals) & !is.na(bioproject) &
                 !is.na(sample_type) & sample_type != "" &
                 !is.na(region) & region != ""
confound_fit  <- lm(fev1_vals[complete_idx] ~ fev1_age_vals[complete_idx] +
                    bioproject[complete_idx] + sample_type[complete_idx] +
                    region[complete_idx])
fev1_adjusted <- residuals(confound_fit) + mean(fev1_vals[complete_idx], na.rm = TRUE)
n_interacting_adj <- n_interacting[complete_idx]
fev1_age_adj      <- fev1_age[complete_idx]

n_boot_val <- 100
n_adj <- length(fev1_adjusted)
boot_lines_adj <- do.call(rbind, lapply(seq_len(n_boot_val), function(b) {
  bi <- sample(seq_len(n_adj), replace = TRUE)
  fit <- lm(fev1_adjusted[bi] ~ n_interacting_adj[bi])
  data.frame(x = range(n_interacting_adj),
             y = coef(fit)[1] + coef(fit)[2] * range(n_interacting_adj), boot = b)
}))

adj_full_lm <- lm(fev1_vals[complete_idx] ~ n_interacting[complete_idx] +
                     fev1_age_vals[complete_idx] + bioproject[complete_idx] +
                     sample_type[complete_idx] + region[complete_idx])
adj_full_s  <- summary(adj_full_lm)
adj_coef <- signif(adj_full_s$coefficients["n_interacting[complete_idx]", "Estimate"], 2)
adj_r2   <- signif(adj_full_s$r.squared, 2)
adj_p    <- format(adj_full_s$coefficients["n_interacting[complete_idx]", "Pr(>|t|)"],
                   digits = 2, scientific = TRUE)
adj_stat_label <- paste0("italic(P)[adj] == ", adj_p,
                         "~~R[adj]^2 == ", adj_r2,
                         "~~beta == '+'*", adj_coef, "*'%'~FEV[1]~Genus^-1")
```

```{r interaction_helpers, include=FALSE}
fit_region_pairs <- function(la) {
  res <- data.frame(genus1 = pairs[1, ], genus2 = pairs[2, ],
                    bx = NA_real_, r = NA_real_, pval = NA_real_,
                    stringsAsFactors = FALSE)
  for (k in seq_len(n_pairs)) {
    y <- la[, pairs[1, k]]; x <- la[, pairs[2, k]]
    if (sd(x) == 0 || sd(y) == 0) next          
    fit <- lm(y ~ x)
    res$bx[k]   <- coef(fit)[2]
    res$r[k]    <- cor(x, y)
    res$pval[k] <- summary(fit)$coefficients[2, 4]
  }
  res
}

ccc <- function(x, y) {
  ok <- is.finite(x) & is.finite(y); x <- x[ok]; y <- y[ok]
  2 * cov(x, y) / (var(x) + var(y) + (mean(x) - mean(y))^2)
}
```

## Alpha Diversity vs Interactions Effect Comparison

```{r alpha_vs_interactions_effect, include=FALSE}
effect_df <- bp_df[bp_complete, ]
effect_df$shannon <- alpha_adj$shannon[match(effect_df$sample, alpha_adj$sample)]
effect_df$n_int   <- n_interacting[match(effect_df$sample, fev1_runs)]
effect_df <- effect_df[complete.cases(
  effect_df[, c("FEV1", "bp_dom", "shannon", "n_int",
                "Age", "BioProject", "Sample_Type")]), ]

effect_metrics <- c(n_int   = "Covarying associations",
                    shannon = "Diversity",
                    bp_dom  = "Dominance")

effect_std_tab <- do.call(rbind, lapply(names(effect_metrics), function(v) {
  f <- as.formula(sprintf("FEV1 ~ scale(%s) + Age + BioProject + Sample_Type", v))
  s <- summary(lm(f, data = effect_df))$coefficients[sprintf("scale(%s)", v), ]
  data.frame(metric    = effect_metrics[[v]], variable = v,
             b_std     = s[["Estimate"]], abs_b_std = abs(s[["Estimate"]]),
             t         = s[["t value"]],  p         = s[["Pr(>|t|)"]],
             stringsAsFactors = FALSE)
}))
effect_std_tab <- effect_std_tab[order(-effect_std_tab$abs_b_std), ]
effect_n <- nrow(effect_df)

# Collinearity: why these are fit separately rather than jointly
effect_cor  <- cor(effect_df[, c("bp_dom", "shannon", "n_int")])
cor_dom_div <- effect_cor["bp_dom", "shannon"]

print(effect_std_tab, row.names = FALSE, digits = 3)
print(round(effect_cor, 3))
sprintf("n = %d; dominance vs diversity r = %.2f", effect_n, cor_dom_div)
```

## Pseudoreplication and repeated sampling

```{r pseudorep_setup, include=FALSE}
reps_meta <- read.csv(
  "rawData/16S_rRNA_gene_amplicon_studies/16S_CF_microbiome_with_reps_long.csv",
  stringsAsFactors = FALSE, check.names = FALSE)
reps_meta$indiv <- paste(reps_meta$BioProject, reps_meta$Sample_ID, sep = "|")

run_indiv <- setNames(reps_meta$indiv, reps_meta$Run)
run_bp    <- setNames(reps_meta$BioProject, reps_meta$Run)
run_reads <- setNames(as.numeric(reps_meta$Bacterial_Read_Count), reps_meta$Run)

runs_per_indiv <- table(reps_meta$indiv)
class_counts   <- table(reps_meta$Sample_Class)
repeated_sampling_summary <- data.frame(
  quantity = c("Sequencing runs", "Unique individuals",
               "Runs from repeat-sampled individuals (%)",
               "Max runs per individual",
               "Class: independent", "Class: longitudinal",
               "Class: replicate", "Class: unresolved"),
  value = c(nrow(reps_meta), length(runs_per_indiv),
            round(100 * sum(runs_per_indiv[reps_meta$indiv] > 1) / nrow(reps_meta), 1),
            max(runs_per_indiv),
            as.integer(class_counts["independent"]), as.integer(class_counts["longitudinal"]),
            as.integer(class_counts["replicate"]),   as.integer(class_counts["unresolved"])),
  stringsAsFactors = FALSE)
dir.create("tables", showWarnings = FALSE)
print(repeated_sampling_summary)

pick_one_per_indiv <- function(runs) {
  runs <- runs[!is.na(run_indiv[runs])]
  runs <- runs[order(run_indiv[runs], -run_reads[runs], runs)]
  runs[!duplicated(run_indiv[runs])]
}

df2A <- data.frame(run = fev1_runs, FEV1 = fev1_vals, n_int = n_interacting,
                   Age = as.numeric(fev1_meta$Age), BioProject = fev1_meta$BioProject,
                   Sample_Type = fev1_meta$Sample_Type, Region = fev1_meta$Region,
                   indiv = run_indiv[fev1_runs], stringsAsFactors = FALSE)

full_b   <- pairwise_results$bx
full_sig <- pairwise_results$significant
concord_2b <- function(b_new) {
  est <- is.finite(b_new)
  c(signs    = 100 * mean(sign(b_new[full_sig & est]) == sign(full_b[full_sig & est])),
    spearman = cor(full_b[est], b_new[est], method = "spearman"),
    ccc      = ccc(full_b[est], b_new[est]))
}
```

```{r pseudorep_mixed, include=FALSE}
ab_mix <- alpha_adj; ab_mix$indiv <- run_indiv[ab_mix$sample]
m1b <- lmer(shannon ~ FEV1_bin + Age + Sample_Type + Region + (1 | BioProject) + (1 | indiv),
            data = ab_mix)
p1B_mixed <- anova(m1b)["FEV1_bin", "Pr(>F)"]

bp_mix <- bp_df; bp_mix$indiv <- run_indiv[bp_mix$sample]
m1c <- lmer(FEV1 ~ bp_dom + Age + Sample_Type + Region + (1 | BioProject) + (1 | indiv),
            data = bp_mix)
c1C_mixed <- summary(m1c)$coefficients["bp_dom", c("Estimate", "Pr(>|t|)")]

m2a <- lmer(FEV1 ~ n_int + Age + Sample_Type + Region + (1 | BioProject) + (1 | indiv),
            data = df2A)
c2A_mixed <- summary(m2a)$coefficients["n_int", c("Estimate", "Pr(>|t|)")]

dir.create("r_objects", showWarnings = FALSE)
pseudo2b_rds <- "r_objects/pseudorep_2B_mixed.rds"
if (file.exists(pseudo2b_rds)) {
  b2b_mixed <- readRDS(pseudo2b_rds)
} else {
  bpv <- run_bp[rownames(log_abund)]; idv <- run_indiv[rownames(log_abund)]
  b2b_mixed <- data.frame(bx = rep(NA_real_, n_pairs), pval = rep(NA_real_, n_pairs))
  for (k in seq_len(n_pairs)) {
    d <- data.frame(y = log_abund[, pairs[1, k]], x = log_abund[, pairs[2, k]],
                    bp = bpv, id = idv)
    m <- tryCatch(suppressWarnings(lmer(y ~ x + (1 | bp) + (1 | id), data = d)),
                  error = function(e) NULL)
    if (!is.null(m)) {
      co <- summary(m)$coefficients
      if ("x" %in% rownames(co)) { b2b_mixed$bx[k] <- co["x", "Estimate"]
                                   b2b_mixed$pval[k] <- co["x", "Pr(>|t|)"] }
    }
  }
  saveRDS(b2b_mixed, pseudo2b_rds)
}
conc2B_mixed <- concord_2b(b2b_mixed$bx)
```

```{r pseudorep_restrict, include=FALSE}
RA_all <- sweep(mat_genus, 1, rowSums(mat_genus), "/") * 100
gg_1A  <- intersect(metacommunity_genera, colnames(RA_all))
RA_all <- RA_all[, gg_1A, drop = FALSE]
prev_f  <- function(M) colSums(M >= 1) / nrow(M) * 100
dom_f   <- function(M) colSums(M >  50) / nrow(M) * 100
conditional_dominance_f <- function(M) { np <- colSums(M >= 1); nd <- colSums(M > 50)
                         ifelse(np == 0, NA_real_, nd / np * 100) }
RA_dd  <- RA_all[pick_one_per_indiv(rownames(RA_all)), , drop = FALSE]
r1A_prev  <- cor(prev_f(RA_all),  prev_f(RA_dd))
r1A_dom   <- cor(dom_f(RA_all),   dom_f(RA_dd))
r1A_conditional_dominance <- cor(conditional_dominance_f(RA_all), conditional_dominance_f(RA_dd), use = "complete.obs")

dd1B <- pick_one_per_indiv(alpha_adj$sample)
p1B_full     <- kruskal.test(shannon_adj ~ FEV1_bin, data = alpha_adj)$p.value
p1B_restrict <- kruskal.test(shannon_adj ~ FEV1_bin,
                             data = alpha_adj[alpha_adj$sample %in% dd1B, ])$p.value

f1C <- FEV1 ~ bp_dom + Age + BioProject + Sample_Type + Region
c1C_full     <- summary(lm(f1C, data = bp_df))$coefficients["bp_dom", c("Estimate", "Pr(>|t|)")]
dd1C <- pick_one_per_indiv(bp_df$sample)
c1C_restrict <- summary(lm(f1C, data = bp_df[bp_df$sample %in% dd1C, ]))$coefficients["bp_dom", c("Estimate", "Pr(>|t|)")]

f2A <- FEV1 ~ n_int + Age + BioProject + Sample_Type + Region
c2A_full     <- summary(lm(f2A, data = df2A))$coefficients["n_int", c("Estimate", "Pr(>|t|)")]
dd2A <- pick_one_per_indiv(df2A$run)
c2A_restrict <- summary(lm(f2A, data = df2A[df2A$run %in% dd2A, ]))$coefficients["n_int", c("Estimate", "Pr(>|t|)")]

dd2B <- pick_one_per_indiv(rownames(log_abund))
rr2B <- fit_region_pairs(log_abund[dd2B, , drop = FALSE])
conc2B_restrict <- concord_2b(rr2B$bx)
```

```{r pseudorep_summary, include=FALSE}
fmt_p <- function(p) {
  if (is.na(p)) return("NA")
  if (p < 1e-15) return("<1e-15")
  if (p < 1e-4)  return(formatC(p, format = "e", digits = 1))
  formatC(p, format = "g", digits = 2)
}

pseudorep_summary <- data.frame(
  Panel = c("1A", "1B", "1C", "2A", "2B"),
  Analysis = c("Prevalence / dominance / conditional dominance per genus",
               "Alpha diversity vs FEV1 bin",
               "Berger-Parker dominance vs FEV1",
               "Interacting genera vs FEV1",
               "Pairwise genus co-occurrence"),
  Full = c("descriptive",
           paste0("KW p=", fmt_p(p1B_full)),
           sprintf("b=%.1f, p=%s", c1C_full["Estimate"], fmt_p(c1C_full["Pr(>|t|)"])),
           sprintf("b=%.2f, p=%s", c2A_full["Estimate"], fmt_p(c2A_full["Pr(>|t|)"])),
           sprintf("%d sig pairs", sum(full_sig))),
  Mixed_model = c("n/a (descriptive)",
                  paste0("FEV1 term p=", fmt_p(p1B_mixed)),
                  sprintf("b=%.1f, p=%s", c1C_mixed["Estimate"], fmt_p(c1C_mixed["Pr(>|t|)"])),
                  sprintf("b=%.2f, p=%s", c2A_mixed["Estimate"], fmt_p(c2A_mixed["Pr(>|t|)"])),
                  sprintf("signs %.0f%%, CCC %.2f", conc2B_mixed["signs"], conc2B_mixed["ccc"])),
  Independent_individuals = c(sprintf("r=%.2f/%.2f/%.2f", r1A_prev, r1A_dom, r1A_conditional_dominance),
                  paste0("KW p=", fmt_p(p1B_restrict)),
                  sprintf("b=%.1f, p=%s", c1C_restrict["Estimate"], fmt_p(c1C_restrict["Pr(>|t|)"])),
                  sprintf("b=%.2f, p=%s", c2A_restrict["Estimate"], fmt_p(c2A_restrict["Pr(>|t|)"])),
                  sprintf("signs %.0f%%, CCC %.2f", conc2B_restrict["signs"], conc2B_restrict["ccc"])),
  stringsAsFactors = FALSE)

dir.create("tables", showWarnings = FALSE)
print(pseudorep_summary)
```

## Tables S2 and S3

```{r tables_S2_S3, include=FALSE}
S2_VARS <- c("Sample_Type", "Age", "Sex", "FEV1")
S2_PERM <- 999
S2_SEED <- 8675309

s2_fit <- function() {
  bm <- as.matrix(bray_dist)
  m  <- cf_microbiome_metadata[match(rownames(bm), cf_microbiome_metadata$Run), ]
  rownames(m) <- rownames(bm)
  for (v in c("Sex", "Sample_Type", "Region", "Instrument"))
    if (!is.null(m[[v]])) m[[v]][!is.na(m[[v]]) & trimws(m[[v]]) == ""] <- NA
  m$Age  <- suppressWarnings(as.numeric(m$Age))
  m$FEV1 <- suppressWarnings(as.numeric(m$FEV1))
  cc  <- function(vs) rownames(m)[Reduce(`&`, lapply(vs, function(v) !is.na(m[[v]])))]
  one <- pick_one_per_indiv(rownames(bm))

  fit <- function(runs, rhs, keep) {
    d <- as.dist(bm[runs, runs]); dat <- m[runs, , drop = FALSE]
    f <- as.formula(paste("d ~", rhs)); environment(f) <- environment()
    set.seed(S2_SEED)
    r <- as.data.frame(adonis2(f, data = dat, permutations = S2_PERM, by = "margin"))
    data.frame(term = keep, n = length(runs), R2 = r[keep, "R2"], p = r[keep, "Pr(>F)"])
  }
  hl <- do.call(rbind, lapply(S2_VARS, function(v) {
    rhs  <- if (v == "Sample_Type") "BioProject + Sample_Type"
            else paste("BioProject + Sample_Type +", v)
    runs <- cc(unique(c("Sample_Type", v)))
    a <- fit(runs, rhs, v); b <- fit(intersect(runs, one), rhs, v)
    data.frame(term = v, n_all = a$n, R2_all = a$R2, p_all = a$p,
               n_one = b$n, R2_one = b$R2, p_one = b$p)
  }))

  ps <- do.call(rbind, unlist(lapply(S2_VARS, function(v)
    lapply(unique(m$BioProject), function(bp) {
      runs <- rownames(m)[m$BioProject == bp & !is.na(m[[v]])]
      if (length(runs) < 20 || length(unique(m[runs, v])) < 2) return(NULL)
      d <- as.dist(bm[runs, runs]); dat <- m[runs, , drop = FALSE]
      f <- as.formula(paste("d ~", v)); environment(f) <- environment()
      set.seed(S2_SEED)
      r <- tryCatch(as.data.frame(adonis2(f, data = dat, permutations = S2_PERM)),
                    error = function(e) NULL)
      if (is.null(r)) return(NULL)
      x <- dat[[v]]
      if (is.numeric(x)) {
        rho <- suppressWarnings(cor(as.vector(d), as.vector(stats::dist(x)),
                                    method = "spearman"))
      } else {
        m2 <- as.matrix(d); same <- outer(x, x, "=="); ut <- upper.tri(m2)
        rho <- mean(m2[ut & !same]) - mean(m2[ut & same])
      }
      data.frame(variable = v, n = length(runs), p = r[1, "Pr(>F)"],
                 direction_rho = rho)
    })), recursive = FALSE))
  list(headline = hl, perstudy = ps)
}

s2_cached <- c("r_objects/permanova_C3b_pervar.rds",
               "r_objects/permanova_E1_one_per_indiv.rds",
               "r_objects/permanova_G_perstudy.rds")
if (all(file.exists(s2_cached))) {
  s2_c3b <- readRDS(s2_cached[1])
  s2_e1  <- readRDS(s2_cached[2]); s2_e1 <- s2_e1[grepl("^E1 C3b", s2_e1$model), ]
  s2_hl  <- do.call(rbind, lapply(S2_VARS, function(v) {
    a <- s2_c3b[s2_c3b$term == v, ][1, ]; b <- s2_e1[s2_e1$term == v, ][1, ]
    data.frame(term = v, n_all = a$n, R2_all = a$R2, p_all = a$p,
               n_one = b$n, R2_one = b$R2, p_one = b$p)
  }))
  s2_ps <- readRDS(s2_cached[3])
} else {
  s2_obj <- s2_fit()
  s2_hl  <- s2_obj$headline
  s2_ps  <- s2_obj$perstudy
  saveRDS(s2_hl, "r_objects/tables_S2_headline.rds")
  saveRDS(s2_ps, "r_objects/tables_S3_perstudy.rds")
}
s2_ps$p_adj <- ave(s2_ps$p, s2_ps$variable, FUN = function(p) p.adjust(p, method = "BH"))

dir.create("tables", showWarnings = FALSE)
table_S2 <- data.frame(
  Variable              = s2_hl$term,
  N_all                 = s2_hl$n_all,
  R2_all                = round(s2_hl$R2_all, 4),
  P_all                 = signif(s2_hl$p_all, 2),
  N_one_per_individual  = s2_hl$n_one,
  R2_one_per_individual = round(s2_hl$R2_one, 4),
  P_one_per_individual  = signif(s2_hl$p_one, 2))
write.csv(table_S2, file.path("tables", "Table_S2_permanova.csv"), row.names = FALSE)

table_S3 <- do.call(rbind, lapply(S2_VARS, function(v) {
  d <- s2_ps[s2_ps$variable == v, ]
  data.frame(Variable = v, Studies = nrow(d),
             `Significant (BH P < 0.05)` = sum(d$p_adj < 0.05, na.rm = TRUE),
             `Fraction agreeing` = round(max(mean(d$direction_rho > 0, na.rm = TRUE),
                                             mean(d$direction_rho < 0, na.rm = TRUE)), 2),
             check.names = FALSE)
}))
write.csv(table_S3, file.path("tables", "Table_S3_perstudy.csv"), row.names = FALSE)

print(table_S2, row.names = FALSE)
print(table_S3, row.names = FALSE)
```

## Process Bioassay Data

```{r process_bioassay_data, include=FALSE}
bioassay_data <- bioassay_data %>%
  mutate(zoi = zoi / 10) %>%
  mutate(inhibition = case_when(
    zoi == 0           ~ 0,
    zoi > 0 & zoi < 1  ~ 1,
    zoi >= 1 & zoi < 2 ~ 2,
    zoi >= 2           ~ 3
  ))

inhibition_counts <- bioassay_data %>%
  count(inhibition) %>%
  mutate(
    inhibition = factor(inhibition, levels = 0:3),
    fraction   = n / sum(n),
    ymax       = cumsum(fraction),
    ymin       = dplyr::lag(ymax, default = 0)
  )

total_assays    <- sum(inhibition_counts$n)
formatted_total <- comma(total_assays)

target_hits <- bioassay_data %>%
  mutate(inhibition = as.integer(inhibition)) %>%
  count(target, inhibition, name = "n") %>%
  group_by(target) %>%
  mutate(
    n_total = sum(n),
    prop    = n / n_total * 100
  ) %>%
  ungroup() %>%
  filter(inhibition %in% 1:3) %>%
  group_by(target) %>%
  mutate(hit_rate = sum(prop)) %>%
  ungroup() %>%
  mutate(target = case_when(
      target == "Achromobacter xylosoxidans" ~ "*A. xylosoxidans*",
      target == "Burkholderia cenocepacia"   ~ "*B. cenocepacia*",
      target == "Burkholderia cepacia"       ~ "*B. cepacia*",
      target == "Neisseria meningitidis"     ~ "*N. meningitidis*",
      target == "Pseudomonas aeruginosa"     ~ "*P. aeruginosa*",
      target == "Rothia mucilaginosa"        ~ "*R. mucilaginosa*",
      target == "Staphylococcus aureus"      ~ "*S. aureus*",
      target == "Streptococcus oralis"       ~ "*S. oralis*",
      TRUE ~ target
    ),
    target     = forcats::fct_reorder(target, hit_rate, .desc = TRUE),
    inhibition = factor(inhibition, levels = c(1, 2, 3))
  )

isolate_inhibition <- left_join(bioassay_data, cf_isolates, by = "isolate")

isolate_inhibition <- isolate_inhibition %>%
  dplyr::select(isolate, hid, genus, target, inhibition) %>%
  mutate(genus = case_when(
    genus == "Staphylococcus"     ~ "*Staphylococcus*",
    genus == "Pseudomonas"        ~ "*Pseudomonas*",
    genus == "Bacillus"           ~ "*Bacillus*",
    genus == "Enterococcus"       ~ "*Enterococcus*",
    genus == "Lacticaseibacillus" ~ "*Lacticaseibacillus*",
    genus == "Klebsiella"         ~ "*Klebsiella*",
    genus == "Streptococcus"      ~ "*Streptococcus*",
    genus == "Escherichia"        ~ "*Escherichia*",
    genus == "Stenotrophomonas"   ~ "*Stenotrophomonas*",
    genus == "Microbacterium"     ~ "*Microbacterium*",
    TRUE ~ "Other"
  )) %>%
  filter(genus != "Other") %>%
  group_by(genus) %>%
  mutate(N_isolates = n_distinct(isolate)) %>%
  group_by(genus, inhibition) %>%
  summarize(
    count = n(),
    N_isolates = dplyr::first(N_isolates),
    .groups = "drop"
  ) %>%
  group_by(genus) %>%
  mutate(
    inhibition = factor(as.character(inhibition), levels = c("0", "1", "2", "3")),
    percent    = count / sum(count),
    inhib_pct  = sum(percent[inhibition %in% c("1", "2", "3")], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(genus = forcats::fct_reorder(genus, inhib_pct, .desc = TRUE))

total_inhibition_per_isolate <- bioassay_data %>%
  group_by(isolate) %>%
  filter(n_distinct(target) == 8) %>%
  summarize(total_inhibition_score = sum(inhibition, na.rm = TRUE), .groups = "drop")

total_inhibition_per_isolate <- left_join(total_inhibition_per_isolate, cf_isolates, by = "isolate")

total_inhibition_per_isolate <- total_inhibition_per_isolate %>%
  dplyr::select(isolate, hid, genus, total_inhibition_score) %>%
  mutate(genus = case_when(
    genus == "Staphylococcus"     ~ "*Staphylococcus*",
    genus == "Pseudomonas"        ~ "*Pseudomonas*",
    genus == "Bacillus"           ~ "*Bacillus*",
    genus == "Enterococcus"       ~ "*Enterococcus*",
    genus == "Lacticaseibacillus" ~ "*Lacticaseibacillus*",
    genus == "Klebsiella"         ~ "*Klebsiella*",
    genus == "Streptococcus"      ~ "*Streptococcus*",
    genus == "Escherichia"        ~ "*Escherichia*",
    genus == "Stenotrophomonas"   ~ "*Stenotrophomonas*",
    genus == "Microbacterium"     ~ "*Microbacterium*",
    TRUE ~ "Other"
  )) %>%
  filter(genus != "Other") %>%
  group_by(genus) %>%
  mutate(
    .sd_g = sd(total_inhibition_score, na.rm = TRUE),
    .mn_g = mean(total_inhibition_score, na.rm = TRUE),
    z_within_genus = ifelse(.sd_g > 0,
                            (total_inhibition_score - .mn_g) / .sd_g,
                            0)
  ) %>%
  ungroup() %>%
  dplyr::select(-.sd_g, -.mn_g) %>%
  mutate(
    .sd_all = sd(total_inhibition_score, na.rm = TRUE),
    .mn_all = mean(total_inhibition_score, na.rm = TRUE),
    z_overall = ifelse(.sd_all > 0,
                       (total_inhibition_score - .mn_all) / .sd_all,
                       0)
  ) %>%
  dplyr::select(-.sd_all, -.mn_all)

median_order <- total_inhibition_per_isolate %>%
  group_by(genus) %>%
  summarize(med = median(z_overall, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(med)) %>%
  pull(genus) %>%
  as.character()

total_inhibition_per_isolate <- total_inhibition_per_isolate %>%
  mutate(genus = factor(as.character(genus), levels = median_order)) %>%
  filter(!is.na(genus))
```

### Stats: Chi-square by Target

```{r stats_chi_target, include=FALSE}
target_hits_stats <- bioassay_data %>%
  mutate(inhibition = as.integer(inhibition)) %>%
  count(target, inhibition, name = "n") %>%
  mutate(inhibition = factor(as.character(inhibition), levels = c("0", "1", "2", "3")))

chi_target_tab  <- xtabs(n ~ target + inhibition, data = target_hits_stats)
chi_target_test <- chisq.test(chi_target_tab, simulate.p.value = TRUE, B = 100000)

chi_target_df <- (nrow(chi_target_tab) - 1) * (ncol(chi_target_tab) - 1)
sprintf("Target test: X^2 = %.1f, df = %d, p < %.1e",
        unname(chi_target_test$statistic), chi_target_df, chi_target_test$p.value)
```

### Stats: Chi-square by Genus

```{r stats_chi_genus, include=FALSE}
isolate_inhibition_for_chi <- isolate_inhibition %>%
  mutate(inhibition = factor(as.character(inhibition), levels = c("0", "1", "2", "3")))

chi_genus_tab  <- xtabs(count ~ genus + inhibition, data = isolate_inhibition_for_chi)
chi_genus_test <- chisq.test(chi_genus_tab, simulate.p.value = TRUE, B = 100000)

chi_genus_df <- (nrow(chi_genus_tab) - 1) * (ncol(chi_genus_tab) - 1)
sprintf("Genus test: X^2 = %.1f, df = %d, p < %.1e",
        unname(chi_genus_test$statistic), chi_genus_df, chi_genus_test$p.value)
```

### Stats: Fligner-Killeen on Z-Score

```{r stats_fligner_global, include=FALSE}
var_fligner_global <- fligner.test(z_overall ~ genus, data = total_inhibition_per_isolate)
var_fligner_global
```

### Stats: Fligner-Killeen Pairwise + Letters (Holm)

```{r stats_fligner_pairwise, include=FALSE}
fligner_gs    <- levels(factor(total_inhibition_per_isolate$genus))
fligner_pairs <- combn(fligner_gs, 2, simplify = FALSE)

fligner_pairs_df <- do.call(rbind, lapply(fligner_pairs, function(g) {
  d <- subset(total_inhibition_per_isolate, genus %in% g)
  data.frame(g1 = g[1], g2 = g[2],
             p  = fligner.test(z_overall ~ genus, data = d)$p.value)
}))

fligner_pairs_df$padj <- p.adjust(fligner_pairs_df$p, method = "BH")

# named vector "g1-g2" for multcompView
fligner_padj_vec <- fligner_pairs_df$padj
names(fligner_padj_vec) <- paste(fligner_pairs_df$g1, fligner_pairs_df$g2, sep = "-")

fligner_letters <- multcompView::multcompLetters(fligner_padj_vec, threshold = 0.05)
fligner_letters_df <- data.frame(
  genus   = names(fligner_letters$Letters),
  letters = unname(fligner_letters$Letters),
  row.names = NULL
)
```

### Stats: Kruskal-Wallis on Z-Score

```{r stats_kw_global, include=FALSE}
kw_global_test <- kruskal.test(z_overall ~ genus, data = total_inhibition_per_isolate)
kw_global_test
```

### Stats: Pairwise Wilcoxon (Holm) + Letters

```{r stats_kw_pairwise_letters, include=FALSE}
kw_pw_test <- pairwise.wilcox.test(
  x      = total_inhibition_per_isolate$z_overall,
  g      = total_inhibition_per_isolate$genus,
  p.adjust.method = "BH",
  exact  = FALSE,
  na.action = na.omit
)

# convert the triangular adjusted p-value matrix to a named vector "g1-g2"
kw_pmat <- kw_pw_test$p.value
kw_idx  <- which(!is.na(kw_pmat), arr.ind = TRUE)
kw_pvec <- kw_pmat[kw_idx]
names(kw_pvec) <- paste(rownames(kw_pmat)[kw_idx[, "row"]],
                        colnames(kw_pmat)[kw_idx[, "col"]],
                        sep = "-")

kw_letters <- multcompView::multcompLetters(kw_pvec, threshold = 0.05)

genus_order_kw <- levels(total_inhibition_per_isolate$genus)
raw_kw   <- kw_letters$Letters[genus_order_kw]
chars_in_order <- unique(unlist(strsplit(raw_kw, "")))
lmap_kw  <- setNames(letters[seq_along(chars_in_order)], chars_in_order)
kw_letters_df <- data.frame(
  genus   = genus_order_kw,
  letters = sapply(raw_kw, function(l)
    paste(sort(lmap_kw[strsplit(l, "")[[1]]]), collapse = "")),
  row.names = NULL
)
```

## Intrasubject Pseudomonas Strain Variability

```{r pseudomonas_intrasubject_variability, include=FALSE}
pseudomonas_variable_hids <- total_inhibition_per_isolate %>%
  filter(grepl("Pseudomonas", genus, ignore.case = TRUE)) %>%
  group_by(hid) %>%
  filter(n() >= 2) %>%
  summarize(z_spread = max(z_within_genus) - min(z_within_genus), .groups = "drop") %>%
  filter(z_spread >= 2) %>%
  pull(hid)

pseudomonas_variable_isolates <- total_inhibition_per_isolate %>%
  filter(grepl("Pseudomonas", genus, ignore.case = TRUE),
         hid %in% pseudomonas_variable_hids) %>%
  group_by(hid) %>%
  summarize(
    z_spread = max(z_within_genus) - min(z_within_genus),
    isolates = paste(paste0(isolate, " (", round(z_within_genus, 2), ")")[order(-z_within_genus)], collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(z_spread))
```

## LC-MS/MS

```{r fig7_S7B_data, include=FALSE}
lcms_p_cutoff <- 0.01     
lcms_n_perm   <- 9999     
lcms_n_boot   <- 1000     

lcms_host_pal <- c(HID07 = "#7570B3", HID08 = "#fdb462ff", HID16 = "#bebada",
                   HID17 = "#b3de69", HID23 = "#f0e04d")
lcms_inh_pal  <- c(Yes = "#6c8197", No = "#c0cad5")
lcms_val_col  <- circlize::colorRamp2(c(-2, 0, 2), c("#4575B4", "#FFFFFF", "#D73027"))

lcms_norm_iso <- function(x) {
  s <- sub("^X[0-9.]+_", "", x); s <- sub("_[A-Z]$", "", s)
  ifelse(grepl("^NB", s), sprintf("NB%04d", as.integer(sub("^NB0*", "", s))), s)
}
lcms_first <- which(colnames(lcms_diff) == "postHoc") + 1
lcms_scols <- colnames(lcms_diff)[lcms_first:ncol(lcms_diff)]
lcms_ciso  <- lcms_norm_iso(lcms_scols)
lcms_icols <- lcms_scols[lcms_ciso != "BHI"]
lcms_iso   <- lcms_ciso[lcms_ciso != "BHI"]

# significant features (raw intensities) + isolate means
lcms_sig     <- which(lcms_diff$pvalue <= lcms_p_cutoff)
lcms_raw     <- as.matrix(lcms_diff[lcms_sig, lcms_icols])
rownames(lcms_raw) <- lcms_diff$name[lcms_sig]; colnames(lcms_raw) <- lcms_icols
lcms_iso_raw <- t(apply(lcms_raw, 1, function(v) tapply(v, lcms_iso, mean)))

# BHI blank-subtracted copy of the same features
lcms_bcols <- lcms_scols[lcms_ciso == "BHI"]
lcms_bmean <- rowMeans(as.matrix(lcms_diff[lcms_sig, lcms_bcols]))
lcms_sub   <- pmax(lcms_raw - lcms_bmean, 0)

lcms_samp_meta <- tibble(sample = lcms_icols, isolate = lcms_iso) %>%
  left_join(lcms_meta, by = "isolate") %>% as.data.frame()
rownames(lcms_samp_meta) <- lcms_samp_meta$sample
lcms_iso_meta <- lcms_meta %>%
  filter(isolate %in% colnames(lcms_iso_raw)) %>%
  arrange(match(isolate, colnames(lcms_iso_raw))) %>% as.data.frame()
lcms_host_pal <- lcms_host_pal[names(lcms_host_pal) %in% lcms_iso_meta$host]

# per-feature z-scored log2 isolate-mean matrix
lcms_zscore <- function(m) { z <- t(scale(t(m))); z[!is.finite(z)] <- 0; z }
lcms_matz   <- lcms_zscore(log2(lcms_iso_raw + 1))
```

## Figures

```{r ggtheme, include=FALSE}
gg_theme <- theme_classic() +
  theme(
    axis.text.y = element_text(color = "black", size = 14),
    axis.title.y = element_text(color = "black", size = 18),
    axis.text.x = element_text(color = "black", size = 16),
    axis.title.x = element_text(color = "black", size = 18),
    axis.ticks.y = element_line(color = "black", linewidth = 0.25),
    axis.ticks.x = element_line(color = "black", linewidth = 0.25),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 18),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
    plot.background = element_blank(),
    legend.position = "none",
    legend.title = element_blank()
  )

fix_panel_size <- function(p, w = unit(3.129, "in"), h = unit(3.129, "in")) {
  g <- if (inherits(p, "ggplot")) ggplotGrob(p) else p
  idx <- which(g$layout$name == "panel")
  for (i in idx) {
    g$widths[g$layout$l[i]]  <- w
    g$heights[g$layout$t[i]] <- h
  }
  g
}
```

```{r fig1_shared, include=FALSE}
fig1_theme <- theme_classic() +
  theme(axis.text    = element_text(color = "black", size = 10),
        axis.title   = element_text(color = "black", size = 12),
        axis.line    = element_blank(),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 0.25),
        plot.tag     = element_text(size = 12, face = "bold"))

pathogen_palette <- c(
  "Pseudomonas"      = "#e9f5a3",
  "Stenotrophomonas" = "#4b86b8",
  "Burkholderia"     = "#c54a53",
  "Escherichia"      = "#e3754f",
  "Achromobacter"    = "#911c43",
  "Staphylococcus"   = "#7dc0a7",
  "Haemophilus"      = "#b5dca9"
)

fig1_pathogen_order <- c("Pseudomonas", "Stenotrophomonas", "Burkholderia",
                         "Escherichia", "Achromobacter", "Staphylococcus",
                         "Haemophilus")
pathogen_in_table <- fig1_pathogen_order[fig1_pathogen_order %in% metacommunity_genera]
oro_in_table      <- sort(metacommunity_genera[!metacommunity_genera %in% cf_pathogens])
fig1_genus_order  <- c(pathogen_in_table, oro_in_table)

# shared age-group palette (used by Figs 1B and 2A); youngest group is green
age_base_cols <- setNames(c("#009E73", "#CC79A7", "#4575B4", "#56B4E9", "gray60"),
                          c("0–11", "12–17", "18–39", "40+", "NA"))
```

### Figure 1A

```{r fig_1A, include=FALSE}
N_BOOT_1A <- 2000

ra_wide_1A <- per_sample_ra %>%
  filter(taxon %in% metacommunity_genera) %>%
  dplyr::select(sample, taxon, ra) %>%
  pivot_wider(names_from = taxon, values_from = ra, values_fill = 0) %>%
  arrange(sample)
ra_mat_1A <- as.matrix(dplyr::select(ra_wide_1A, -sample))
n_samp_1A <- nrow(ra_mat_1A)

prev_fn_1A  <- function(M) colSums(M >= 1)  / nrow(M) * 100
dom_fn_1A   <- function(M) colSums(M >  50) / nrow(M) * 100
conditional_dominance_fn_1A <- function(M) {
  np <- colSums(M >= 1); nd <- colSums(M > 50)
  ifelse(np == 0, NA_real_, nd / np * 100)
}

boot_arr_1A <- array(
  NA_real_,
  dim = c(N_BOOT_1A, ncol(ra_mat_1A), 3),
  dimnames = list(NULL, colnames(ra_mat_1A),
                  c("prev_1_pct", "dominant_pct", "conditional_dominance"))
)
for (b in seq_len(N_BOOT_1A)) {
  Mb <- ra_mat_1A[sample.int(n_samp_1A, n_samp_1A, replace = TRUE), , drop = FALSE]
  boot_arr_1A[b, , "prev_1_pct"]       <- prev_fn_1A(Mb)
  boot_arr_1A[b, , "dominant_pct"]     <- dom_fn_1A(Mb)
  boot_arr_1A[b, , "conditional_dominance"] <- conditional_dominance_fn_1A(Mb)
}
ci_lo_1A <- apply(boot_arr_1A, c(2, 3), quantile, probs = 0.025, na.rm = TRUE)
ci_hi_1A <- apply(boot_arr_1A, c(2, 3), quantile, probs = 0.975, na.rm = TRUE)

ci_long_1A <- bind_rows(
  tibble(genus = rownames(ci_lo_1A), metric = "prev_1_pct",
         lo = ci_lo_1A[, "prev_1_pct"], hi = ci_hi_1A[, "prev_1_pct"]),
  tibble(genus = rownames(ci_lo_1A), metric = "dominant_pct",
         lo = ci_lo_1A[, "dominant_pct"], hi = ci_hi_1A[, "dominant_pct"]),
  tibble(genus = rownames(ci_lo_1A), metric = "conditional_dominance",
         lo = ci_lo_1A[, "conditional_dominance"], hi = ci_hi_1A[, "conditional_dominance"])
)

metric_cols <- c(
  "Prevalence (\u22651% Relative Abundance)" = "#c0cad5",
  "Dominance (>50% Relative Abundance)"      = "#6c8197",
  "Conditional Dominance (Dominance/Prevalence)"        = "#94a8bf"
)

plot_df_1A <- metacommunity_table %>%
  dplyr::select(genus, prev_1_pct, dominant_pct, conditional_dominance) %>%
  pivot_longer(cols = c(prev_1_pct, dominant_pct, conditional_dominance),
               names_to = "metric", values_to = "value") %>%
  left_join(ci_long_1A, by = c("genus", "metric")) %>%
  mutate(
    metric = recode(metric,
                    prev_1_pct       = "Prevalence (\u22651% Relative Abundance)",
                    dominant_pct     = "Dominance (>50% Relative Abundance)",
                    conditional_dominance = "Conditional Dominance (Dominance/Prevalence)"),
    metric = factor(metric, levels = c("Prevalence (\u22651% Relative Abundance)",
                                       "Dominance (>50% Relative Abundance)",
                                       "Conditional Dominance (Dominance/Prevalence)")),
    genus  = factor(genus, levels = fig1_genus_order)
  )

genus_labels_1A <- setNames(
  sapply(fig1_genus_order, function(g) {
    if (g %in% cf_pathogens) {
      paste0("<i><span style='color:#E69F00; font-weight:bold;'>", g, "</span></i>")
    } else {
      paste0("<i>", g, "</i>")
    }
  }),
  fig1_genus_order
)

panel_1a <- ggplot(plot_df_1A, aes(x = genus, y = value, fill = metric)) +
  geom_hline(yintercept = c(20, 40, 60, 80),
             color = "gray85", linewidth = 0.3) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75,
           color = "black", linewidth = 0.2) +
  geom_errorbar(aes(ymin = lo, ymax = hi),
                position = position_dodge(width = 0.8),
                width = 0.3, linewidth = 0.3, color = "black", na.rm = TRUE) +
  scale_fill_manual(values = metric_cols, name = NULL) +
  scale_x_discrete(labels = genus_labels_1A) +
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80),
                     expand = expansion(mult = c(0, 0.05))) +
  labs(x = NULL, y = "Percent (%)") +
  fig1_theme +
  theme(axis.text.x          = ggtext::element_markdown(angle = 45, hjust = 1, size = 10),
        legend.position      = c(0.005, 0.97),
        legend.justification  = c(0, 1),
        legend.background    = element_rect(fill = "white", color = "black",
                                              linewidth = 0.25),
        legend.margin        = margin(2, 4, 2, 4),
        legend.text          = element_text(size = 10),
        legend.key.size      = unit(0.4, "cm"))
```

### Figure 1B

```{r fig_1B, include=FALSE}
set.seed(8675309)

alpha_cont <- alpha_adj
alpha_cont$FEV1 <- as.numeric(alpha_cont$FEV1)
alpha_cont$Age  <- as.numeric(alpha_cont$Age)
alpha_cont <- alpha_cont[!is.na(alpha_cont$FEV1) &
                         !is.na(alpha_cont$shannon) &
                         !is.na(alpha_cont$Age), ]

alpha_cont$FEV1_adj <- residuals(
  lm(FEV1 ~ Age + BioProject + Sample_Type + Region, data = alpha_cont)) +
  mean(alpha_cont$FEV1)

alpha_cont$age_bin <- cut(alpha_cont$Age, breaks = c(-Inf, 12, 18, 40, Inf),
                          right = FALSE, labels = c("0\u201311", "12\u201317", "18\u201339", "40+"))
age_bins_1b <- levels(alpha_cont$age_bin)[table(alpha_cont$age_bin) > 0]
age_n_1b    <- table(alpha_cont$age_bin)[age_bins_1b]
age_lab_1b  <- setNames(paste0(age_bins_1b, " (N=", age_n_1b, ")"), age_bins_1b)
alpha_cont$age_group <- factor(age_lab_1b[as.character(alpha_cont$age_bin)],
                               levels = age_lab_1b[age_bins_1b])
cols_1b <- setNames(unname(age_base_cols[age_bins_1b]), age_lab_1b[age_bins_1b])

n_boot_1b <- 100; n_1b <- nrow(alpha_cont)
boot_lines_1b <- do.call(rbind, lapply(seq_len(n_boot_1b), function(b) {
  bi  <- sample.int(n_1b, n_1b, replace = TRUE)
  fit <- lm(alpha_cont$FEV1_adj[bi] ~ alpha_cont$shannon[bi])
  data.frame(x = range(alpha_cont$shannon),
             y = coef(fit)[1] + coef(fit)[2] * range(alpha_cont$shannon), boot = b)
}))

panel_1b <- ggplot(alpha_cont, aes(x = shannon, y = FEV1_adj)) +
  geom_line(data = boot_lines_1b, aes(x = x, y = y, group = boot),
            color = "lightblue", alpha = 0.15) +
  geom_point(aes(color = age_group), alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", color = "#4575B4", se = FALSE, linewidth = 1.1) +
  scale_color_manual(values = cols_1b, name = "Age") +
  labs(x = expression(Diversity~"("*italic(H)*"'"*")"),
       y = expression(Lung~Function~"("*FEV[1]~pp[Adj]*")")) +
  fig1_theme +
  theme(legend.position = "bottom",
        legend.text  = element_text(size = 10),
        legend.title = element_text(size = 10, face = "bold")) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE,
                              override.aes = list(alpha = 1, size = 2)))
```

### Figure 1C

```{r fig_1C, include=FALSE}
bp_plot_df <- bp_adj_df %>%
  mutate(dom_type = factor(if_else(dominant_pathogen == "Other",
                                   "Oropharyngeal", "Pathogen"),
                           levels = c("Pathogen", "Oropharyngeal")))

dom_type_counts <- bp_plot_df %>% count(dom_type)
dom_type_labels <- setNames(
  vapply(levels(bp_plot_df$dom_type), function(g) {
    n <- dom_type_counts$n[dom_type_counts$dom_type == g]
    if (length(n) == 0) n <- 0
    paste0(g, " (N=", scales::comma(n), ")")
  }, character(1)),
  levels(bp_plot_df$dom_type)
)
dom_type_cols <- c("Pathogen" = "#E69F00", "Oropharyngeal" = "gray60")

panel_1c <- ggplot(bp_plot_df, aes(x = bp_dom, y = FEV1_adj)) +
  geom_line(data = boot_lines_bp, aes(x = x, y = y, group = boot),
            color = "lightpink", alpha = 0.15) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_jitter(aes(color = dom_type),
              width = 0.005, alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", color = "#D73027", se = FALSE, linewidth = 1.1) +
  scale_color_manual(values = dom_type_cols, labels = dom_type_labels,
                     name = "Dominant
Taxon") +
  scale_x_continuous(breaks = seq(0.2, 1.0, 0.2)) +
  scale_y_continuous() +
  labs(x = "Dominance (BP)",
       y = expression(Lung~Function~"("*FEV[1]~pp[Adj]*")")) +
  fig1_theme +
  theme(legend.position    = "bottom",
        legend.text        = element_text(size = 9),
        legend.title       = element_text(size = 9, face = "bold"),
        legend.key.height  = unit(0.35, "cm"),
        legend.spacing.y   = unit(0, "cm")) +
  guides(color = guide_legend(nrow = 1, reverse = TRUE,
                               override.aes = list(alpha = 1, size = 1.5)))
```

### Figure 1

```{r fig_1, include=FALSE}
fig2_panel_size <- grid::unit(3.2, "in")

pad_to_height <- function(g, target_h_in) {
  cur_h <- sum(grid::convertHeight(g$heights, "in", valueOnly = TRUE))
  if (cur_h < target_h_in) {
    g <- gtable::gtable_add_rows(g, grid::unit(target_h_in - cur_h, "in"))
  }
  g
}

leg_1b <- cowplot::get_legend(panel_1b)
leg_1c <- cowplot::get_legend(panel_1c)

fix_panel_size_grob <- function(g, w, h) {
  panels <- g$layout[g$layout$name == "panel", ]
  l_idx <- panels$l[1]; r_idx <- panels$r[1]
  t_idx <- panels$t[1]; b_idx <- panels$b[1]
  g$widths[l_idx:r_idx]  <- w
  g$heights[t_idx:b_idx] <- h
  g
}

panel_1b_gg <- panel_1b + labs(tag = "B") +
  theme(legend.position = "none", aspect.ratio = NULL,
        plot.tag = element_text(size = 12, face = "bold"))

panel_1c_gg <- panel_1c + labs(tag = "C") +
  theme(legend.position = "none", aspect.ratio = NULL,
        plot.tag = element_text(size = 12, face = "bold"))

panel_1a_gg <- panel_1a + labs(tag = "A") +
  theme(plot.tag    = element_text(size = 12, face = "bold"),
        plot.margin = margin(b = 30, unit = "pt"))

al_lr        <- cowplot::align_plots(panel_1a_gg, panel_1b_gg, panel_1c_gg,
                                      align = "v", axis = "lr")
panel_1a_al  <- al_lr[[1]]

panel_1b_sized <- fix_panel_size_grob(al_lr[[2]], fig2_panel_size, fig2_panel_size)
panel_1c_sized <- fix_panel_size_grob(al_lr[[3]], fig2_panel_size, fig2_panel_size)

bc_heights      <- c(sum(grid::convertHeight(panel_1b_sized$heights, "in", valueOnly = TRUE)),
                     sum(grid::convertHeight(panel_1c_sized$heights, "in", valueOnly = TRUE)))
bc_hmax         <- max(bc_heights)
panel_1b_padded <- pad_to_height(panel_1b_sized, bc_hmax)
panel_1c_padded <- pad_to_height(panel_1c_sized, bc_hmax)

bc_top  <- cowplot::plot_grid(panel_1b_padded, panel_1c_padded, ncol = 2)
bc_legs <- cowplot::plot_grid(leg_1b, leg_1c, ncol = 2)
bc_row  <- cowplot::plot_grid(bc_top, bc_legs, ncol = 1, rel_heights = c(1, 0.26))

figure_1 <- cowplot::plot_grid(panel_1a_al, bc_row, ncol = 1, rel_heights = c(1, 1.2))

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_1.pdf", figure_1, units = "in", width = 9, height = 9, device = cairo_pdf)
```

### Figure 2A

```{r fig_2A, include=FALSE}
bx_mat   <- matrix(0, n_genera, n_genera, dimnames = list(genera, genera))
pval_mat <- matrix(1, n_genera, n_genera, dimnames = list(genera, genera))

for (i in seq_len(nrow(pairwise_results))) {
  g1 <- pairwise_results$genus1[i]
  g2 <- pairwise_results$genus2[i]
  bx_mat[g1, g2]  <- bx_mat[g2, g1]  <- pairwise_results$bx[i]
  pval_mat[g1, g2] <- pval_mat[g2, g1] <- pairwise_results$pval[i]
}

pathogen_order  <- c("Pseudomonas", "Stenotrophomonas", "Burkholderia",
                     "Escherichia", "Achromobacter", "Staphylococcus", "Haemophilus")
pathogen_genera <- pathogen_order[pathogen_order %in% genera]
anaerobe_genera <- sort(genera[!genera %in% cf_pathogens])
genus_order     <- c(pathogen_genera, anaerobe_genera)

isolate_genera  <- unique(cf_isolates$genus)
target_genera   <- unique(sub(" .*", "", bioassay_data$target))
bioassay_genera <- unique(c(isolate_genera, target_genera))

axis_colors <- ifelse(genus_order %in% cf_pathogens, "#E69F00", "#000000")
axis_labels  <- sapply(seq_along(genus_order), function(i) {
  g <- genus_order[i]; col <- axis_colors[i]
  if (g %in% bioassay_genera)
    paste0("<span style='color:", col, "'><b><i>", g, "</i></b></span>")
  else
    paste0("<span style='color:", col, "'><i>", g, "</i></span>")
})
names(axis_labels) <- genus_order

bubble_df <- pairwise_results %>%
  mutate(direction = ifelse(bx > 0, "Positive", "Negative"),
         genus1 = factor(genus1, levels = genus_order),
         genus2 = factor(genus2, levels = genus_order)) %>%
  filter(significant) %>%
  mutate(r = as.integer(genus1), c = as.integer(genus2),
         row_genus = ifelse(r > c, as.character(genus1), as.character(genus2)),
         col_genus = ifelse(r > c, as.character(genus2), as.character(genus1))) %>%
  mutate(row_genus = factor(row_genus, levels = genus_order),
         col_genus = factor(col_genus, levels = genus_order)) %>%
  dplyr::select(row_genus, col_genus, bx, direction)

n_go    <- length(genus_order)
s       <- seq_len(n_go - 1)
h_lines <- data.frame(y = s, xmin = 0.5, xmax = (n_go - s) + 1.0)
v_lines <- data.frame(x = s, ymin = 0.5, ymax = (n_go - s) + 1.0)

panel_2a <- ggplot(bubble_df, aes(x = col_genus, y = row_genus)) +
  geom_segment(data = h_lines, aes(x = xmin, xend = xmax, y = y, yend = y),
               color = "gray80", linewidth = 0.15, inherit.aes = FALSE) +
  geom_segment(data = v_lines, aes(x = x, xend = x, y = ymin, yend = ymax),
               color = "gray80", linewidth = 0.15, inherit.aes = FALSE) +
  geom_point(aes(size = abs(bx), fill = direction), shape = 21,
             color = "black", stroke = 0.3) +
  scale_fill_manual(values = c("Negative" = "#D73027", "Positive" = "#4575B4"),
                    name = "Association") +
  scale_size_continuous(range = c(0.75, 5.25), name = "Coefficient Magnitude",
                        breaks = c(0.05, 0.1, 0.5, 0.75),
                        labels = c("0.05", "0.10", "0.50", "0.75")) +
  scale_x_discrete(limits = genus_order[-length(genus_order)],
                   labels = axis_labels, drop = FALSE) +
  scale_y_discrete(limits = rev(genus_order[-1]),
                   labels = axis_labels[rev(genus_order[-1])], drop = FALSE) +
  labs(x = NULL, y = NULL, tag = "A") +
  gg_theme +
  theme(axis.line = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
        axis.text.x  = ggtext::element_markdown(angle = 45, hjust = 1, size = 10),
        axis.text.y  = ggtext::element_markdown(size = 10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        legend.position = "bottom",
        legend.text  = element_text(size = 10),
        legend.title = element_text(size = 10, face = "bold"),
        legend.box = "vertical",
        panel.grid.major = element_blank(),
        plot.tag = element_text(size = 12, face = "bold")) +
  guides(size = guide_legend(nrow = 2, byrow = TRUE))
```

### Figure 2A inset

```{r fig_2A_upset_inset, include=FALSE}
sig_pairs_upset <- pairwise_results %>%
  filter(significant) %>%
  mutate(
    g1_path       = genus1 %in% cf_pathogens,
    g2_path       = genus2 %in% cf_pathogens,
    direction     = factor(ifelse(bx > 0, "Positive", "Negative"),
                           levels = c("Positive", "Negative")),
    Pathogen      = g1_path | g2_path,
    Oropharyngeal = !g1_path | !g2_path
  )

direction_cols_upset <- c("Negative" = "#D73027", "Positive" = "#4575B4")
pathogen_gold_upset  <- "#E69F00"

inset_overrides_2b <- theme(
  axis.text.y  = element_text(color = "black", size = 10),
  axis.title.y = element_text(color = "black", size = 10),
  axis.text.x  = element_text(color = "black", size = 10),
  axis.title.x = element_text(color = "black", size = 10),
  axis.text    = element_text(size = 10),
  strip.text   = element_text(size = 10),
  legend.position = "none",
  plot.background = element_rect(fill = "white", color = NA),
  plot.margin     = margin(2, 3, 2, 2)
)

upset_inset_2a <- upset(
  data         = as.data.frame(sig_pairs_upset),
  intersect    = c("Oropharyngeal", "Pathogen"),
  name         = NULL,
  sort_sets          = FALSE,
  sort_intersections = FALSE,
  intersections = list(
    "Pathogen",
    c("Pathogen", "Oropharyngeal"),
    "Oropharyngeal"
  ),
  stripes      = upset_stripes(colors = c("white", "white")),
  width_ratio  = 0.22,
  height_ratio = 0.55,
  queries = list(
    upset_query(set   = "Pathogen",
                fill  = pathogen_gold_upset,
                color = pathogen_gold_upset,
                only_components = c("intersections_matrix"))
  ),
  base_annotations = list(
    "# of Pairs" = intersection_size(
      counts  = FALSE,
      mapping = aes(fill = direction),
      width   = 0.6
    ) +
      scale_fill_manual(values = direction_cols_upset, guide = "none") +
      gg_theme + inset_overrides_2b +
      theme(panel.border = element_blank(),
            axis.line.y  = element_line(color = "black", linewidth = 0.25),
            axis.line.x  = element_line(color = "black", linewidth = 0.25),
            axis.text.x  = element_blank(),
            axis.title.x = element_blank(),
            axis.ticks.x = element_blank())
  ),
  set_sizes = (
    upset_set_size(geom = geom_bar(width = 0.6)) +
      gg_theme + inset_overrides_2b +
      theme(panel.border = element_blank(),
            axis.line.x  = element_line(color = "black", linewidth = 0.25),
            axis.text.x  = element_text(
              color = c("black", "transparent", "black",
                        "transparent", "black"),
              angle = 45, hjust = 1, size = 10
            ),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y  = element_blank(),
            axis.ticks.y = element_blank())
  ),
  matrix = intersection_matrix(geom = geom_point(size = 2.8)),
  themes = upset_modify_themes(list(
    "intersections_matrix" =
      gg_theme + inset_overrides_2b +
      theme(axis.text.y  = element_blank(),
            axis.text.x  = element_blank(),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            axis.ticks   = element_blank(),
            panel.border = element_blank(),
            panel.background = element_blank())
  ))
)

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_2A_upset_inset.pdf", upset_inset_2a,
       width = 2.2, height = 1.95, device = cairo_pdf)
```

### Figure 2B

```{r fig_2B, include=FALSE}
# dominance is taken from bp_df so the definition matches Figure 1C exactly
b2_runs <- fev1_runs[complete_idx]
b2_dom  <- setNames(bp_df$dominant_pathogen, bp_df$sample)[b2_runs]
b2_type <- factor(ifelse(is.na(b2_dom) | b2_dom == "Other",
                         "Oropharyngeal", "Pathogen"),
                  levels = c("Pathogen", "Oropharyngeal"))
b2_cnt  <- table(b2_type)
b2_labs <- setNames(paste0(names(b2_cnt), " (N=",
                           scales::comma(as.integer(b2_cnt)), ")"),
                    names(b2_cnt))

plot_df_adj <- data.frame(
  n_taxa    = n_interacting_adj,
  fev1_adj  = fev1_adjusted,
  dom_type  = b2_type)

panel_2b <- ggplot(plot_df_adj, aes(x = n_taxa, y = fev1_adj)) +
  geom_line(data = boot_lines_adj, aes(x = x, y = y, group = boot),
            color = "lightblue", alpha = 0.15) +
  geom_jitter(aes(color = dom_type), width = 0.4, alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", color = "#4575B4", se = FALSE, linewidth = 1.1) +
  scale_color_manual(values = dom_type_cols, labels = b2_labs,
                     name = "Dominant
Taxon") +
  scale_y_continuous(breaks = seq(25, 125, 25)) +
  coord_cartesian(ylim = c(12, 141)) +
  labs(x = "Pairs of Covarying Genera",
       y = expression(Lung~Function~"("*FEV[1]~pp[Adj]*")"),
       tag = "B") +
  gg_theme +
  theme(axis.line = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
        axis.text.x  = element_text(size = 10, color = "black"),
        axis.text.y  = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        legend.position = "bottom",
        legend.text  = element_text(size = 10),
        legend.title = element_text(size = 10, face = "bold"),
        plot.tag = element_text(size = 12, face = "bold")) +
  guides(color = guide_legend(nrow = 1, reverse = TRUE,
                              override.aes = list(alpha = 1, size = 2.5)))
```

### Figure 2C

```{r fig_2C, include=FALSE}
C2_DEPTH  <- 1000    
C2_DRAWS  <- 2000   
C2_BREAKS <- c(1, 10, 20, 35, 50, 75)
C2_LABS   <- c(">0-1", "1-10", "10-20", "20-35", "35-50", "50-75", ">75")
C2_POS    <- c(0.5, 5.5, 15, 27.5, 42.5, 62.5, 87.5)   # bin midpoints, linear scale

c2_oro   <- setdiff(colnames(genus_matrix_filt), cf_pathogens)
c2_ra    <- 100 * genus_matrix_filt / rowSums(genus_matrix_filt)
c2_pRA   <- rowSums(c2_ra[, cf_pathogens, drop = FALSE])
c2_pRds  <- rowSums(genus_matrix_filt[, cf_pathogens, drop = FALSE])
c2_pRA_all <- c2_pRA           # kept for the lung-function series, which is not
c2_pRds_all <- c2_pRds         # depth-filtered
c2_keep  <- rowSums(genus_matrix_filt[, c2_oro, drop = FALSE]) >= C2_DEPTH
c2_O     <- genus_matrix_filt[c2_keep, c2_oro, drop = FALSE]
c2_P     <- c2_O / rowSums(c2_O)
c2_pRA   <- c2_pRA[c2_keep]; c2_pRds <- c2_pRds[c2_keep]

c2_bin <- ifelse(c2_pRds == 0, "0",
           as.character(cut(c2_pRA, c(-Inf, C2_BREAKS, Inf), labels = C2_LABS)))
c2_bin <- factor(c2_bin, levels = c("0", C2_LABS))
c2_idx <- lapply(levels(c2_bin), function(l) which(c2_bin == l))

c2_pairs <- combn(c2_oro, 2); c2_i1 <- c2_pairs[1, ]; c2_i2 <- c2_pairs[2, ]
c2_bvec  <- function(M) {
  L <- log10(M + 1); C <- cov(L); C[cbind(c2_i1, c2_i2)] / diag(C)[c2_i2]
}

c2_rar <- function(Pm, D) {
  n <- nrow(Pm); K <- ncol(Pm); X <- matrix(0L, n, K)
  rem <- rep(D, n); remp <- rep(1, n)
  for (j in seq_len(K - 1)) {
    q <- pmin(pmax(Pm[, j] / remp, 0), 1); q[!is.finite(q)] <- 0
    x <- rbinom(n, rem, q); X[, j] <- x; rem <- rem - x; remp <- remp - Pm[, j]
  }
  X[, K] <- rem; colnames(X) <- colnames(Pm); X
}

dir.create("r_objects", showWarnings = FALSE)
if (file.exists("r_objects/fig2C_boot.rds")) {
  c2_mb <- readRDS("r_objects/fig2C_boot.rds")
} else {
  set.seed(97)
  c2_mb <- matrix(NA_real_, C2_DRAWS, length(levels(c2_bin)))
  for (b in seq_len(C2_DRAWS)) for (j in seq_along(c2_idx)) {
    s <- sample(c2_idx[[j]], length(c2_idx[[j]]), replace = TRUE)
    c2_mb[b, j] <- mean(c2_bvec(c2_rar(c2_P[s, , drop = FALSE], C2_DEPTH)))
  }
  colnames(c2_mb) <- levels(c2_bin)
  saveRDS(c2_mb, "r_objects/fig2C_boot.rds")
}

c2_rel <- c2_mb / c2_mb[, 1]
c2_pl  <- 2:ncol(c2_mb)
c2_bp  <- function(v) { p1 <- mean(v <= 0); max(2 * min(p1, 1 - p1), 1 / C2_DRAWS) }
c2_cmb <- combn(length(c2_pl), 2)
c2_Q   <- p.adjust(vapply(seq_len(ncol(c2_cmb)), function(k)
  c2_bp(c2_mb[, c2_pl][, c2_cmb[1, k]] - c2_mb[, c2_pl][, c2_cmb[2, k]]),
  numeric(1)), "BH")

c2_M <- matrix(1, length(c2_pl), length(c2_pl), dimnames = list(C2_LABS, C2_LABS))
for (k in seq_len(ncol(c2_cmb)))
  c2_M[c2_cmb[1, k], c2_cmb[2, k]] <- c2_M[c2_cmb[2, k], c2_cmb[1, k]] <- c2_Q[k]

c2_adj <- c2_M > 0.05; diag(c2_adj) <- TRUE
c2_mn  <- colMeans(c2_mb[, c2_pl]); c2_cl <- list()
for (sz in length(c2_pl):1) for (cb in asplit(combn(length(c2_pl), sz), 2)) {
  v <- as.integer(cb)
  if (all(c2_adj[v, v]) && !any(vapply(c2_cl, function(c0) all(v %in% c0), logical(1))))
    c2_cl[[length(c2_cl) + 1]] <- v
}
c2_cl  <- c2_cl[order(vapply(c2_cl, function(v) -max(c2_mn[v]), numeric(1)))]
c2_ltr <- setNames(rep("", length(c2_pl)), C2_LABS)
for (i in seq_along(c2_cl)) for (j in c2_cl[[i]])
  c2_ltr[C2_LABS[j]] <- paste0(c2_ltr[C2_LABS[j]], letters[i])

fig2c_df <- data.frame(
  stratum = C2_LABS, pos = C2_POS,
  n       = as.integer(table(c2_bin))[c2_pl],
  rel     = colMeans(c2_rel[, c2_pl]),
  lo      = apply(c2_rel[, c2_pl], 2, quantile, 0.025),
  hi      = apply(c2_rel[, c2_pl], 2, quantile, 0.975),
  letter  = c2_ltr[C2_LABS], stringsAsFactors = FALSE)

C2_FEV_CUT <- 80
c2_meta  <- cf_microbiome_metadata[match(rownames(genus_matrix_filt),
                                         cf_microbiome_metadata$Run), ]
c2_fev   <- suppressWarnings(as.numeric(c2_meta$FEV1))
c2_binF  <- ifelse(c2_pRds_all == 0, "0",
             as.character(cut(c2_pRA_all, c(-Inf, C2_BREAKS, Inf), labels = C2_LABS)))
c2_binF  <- factor(c2_binF, levels = c("0", C2_LABS))
c2_hi    <- !is.na(c2_fev) & c2_fev >= C2_FEV_CUT
c2_okF   <- !is.na(c2_fev) & c2_binF != "0"
c2_idxF  <- lapply(C2_LABS, function(l) which(c2_binF == l & c2_okF))

if (file.exists("r_objects/fig2C_lf_boot.rds")) {
  c2_fb <- readRDS("r_objects/fig2C_lf_boot.rds")
} else {
  set.seed(11)
  c2_fb <- matrix(NA_real_, C2_DRAWS, length(C2_LABS))
  for (b in seq_len(C2_DRAWS)) for (j in seq_along(c2_idxF))
    c2_fb[b, j] <- 100 * mean(c2_hi[sample(c2_idxF[[j]],
                                           length(c2_idxF[[j]]), replace = TRUE)])
  saveRDS(c2_fb, "r_objects/fig2C_lf_boot.rds")
}
colnames(c2_fb) <- C2_LABS

c2_QF <- p.adjust(vapply(seq_len(ncol(c2_cmb)), function(k)
  c2_bp(c2_fb[, c2_cmb[1, k]] - c2_fb[, c2_cmb[2, k]]), numeric(1)), "BH")
c2_MF <- matrix(1, length(C2_LABS), length(C2_LABS),
                dimnames = list(C2_LABS, C2_LABS))
for (k in seq_len(ncol(c2_cmb)))
  c2_MF[c2_cmb[1, k], c2_cmb[2, k]] <- c2_MF[c2_cmb[2, k], c2_cmb[1, k]] <- c2_QF[k]
c2_adjF <- c2_MF > 0.05; diag(c2_adjF) <- TRUE
c2_mnF  <- colMeans(c2_fb); c2_clF <- list()
for (sz in length(C2_LABS):1) for (cb in asplit(combn(length(C2_LABS), sz), 2)) {
  v <- as.integer(cb)
  if (all(c2_adjF[v, v]) && !any(vapply(c2_clF, function(c0) all(v %in% c0), logical(1))))
    c2_clF[[length(c2_clF) + 1]] <- v
}
c2_clF  <- c2_clF[order(vapply(c2_clF, function(v) -max(c2_mnF[v]), numeric(1)))]
c2_ltrF <- setNames(rep("", length(C2_LABS)), C2_LABS)
for (i in seq_along(c2_clF)) for (j in c2_clF[[i]])
  c2_ltrF[C2_LABS[j]] <- paste0(c2_ltrF[C2_LABS[j]], letters[i])

fig2c_lf <- data.frame(stratum = C2_LABS, pos = C2_POS,
  n = lengths(c2_idxF), pct = c2_mnF,
  lo = apply(c2_fb, 2, quantile, 0.025),
  hi = apply(c2_fb, 2, quantile, 0.975),
  letter = c2_ltrF[C2_LABS], stringsAsFactors = FALSE)

c2_track <- vapply(seq_len(C2_DRAWS), function(b)
  cor(c2_rel[b, c2_pl], c2_fb[b, ], method = "spearman"), numeric(1))
c2_track_p <- max(2 * min(mean(c2_track <= 0), mean(c2_track >= 0)), 1 / C2_DRAWS)

C2_Y  <- c(0.12, 1.41); C2_F <- c(0, 85)
c2_f2y <- function(v) C2_Y[1] + (v - C2_F[1]) * diff(C2_Y) / diff(C2_F)
C2_OFF <- 1.6      
c2_a <- transform(fig2c_df, x = pos - C2_OFF)
c2_b <- transform(fig2c_lf,
  x = pos + C2_OFF, y = c2_f2y(pct), ylo = c2_f2y(lo), yhi = c2_f2y(hi))

panel_2c <- ggplot() +
  geom_errorbar(data = c2_a, aes(x, ymin = lo, ymax = hi), width = 1.8,
                linewidth = 0.55, color = "black") +
  geom_point(data = c2_a, aes(x, rel), size = 3.0, color = "black") +
  geom_text(data = c2_a, aes(x, hi + 0.05, label = letter),
            size = 3.5, color = "black") +
  geom_errorbar(data = c2_b, aes(x, ymin = ylo, ymax = yhi), width = 1.8,
                linewidth = 0.55, color = "#4575B4") +
  geom_point(data = c2_b, aes(x, y), size = 3.0, color = "#4575B4") +
  geom_text(data = c2_b, aes(x, yhi + 0.05, label = letter),
            size = 3.5, color = "#4575B4") +
  scale_x_continuous("Pathogen Relative Abundance (%)",
                     breaks = C2_POS, labels = C2_LABS) +
  scale_y_continuous("Relative Mean Oropharyngeal Covariation",
    breaks = seq(0.25, 1.25, 0.25),
    labels = sprintf("%.2f", seq(0.25, 1.25, 0.25)),
    sec.axis = sec_axis(~ C2_F[1] + (. - C2_Y[1]) * diff(C2_F) / diff(C2_Y),
      name = expression("High Lung Function ("*FEV[1]~pp>=80*", %)"),
      breaks = seq(0, 80, 20))) +
  coord_cartesian(xlim = c(-1.5, 90), ylim = C2_Y) +
  labs(tag = "C") +
  gg_theme +
  theme(axis.line    = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
        axis.text.x  = element_text(size = 10, color = "black", angle = 45, hjust = 1),
        axis.title.x      = element_text(size = 12, color = "black"),
        axis.text.y.left  = element_text(size = 10, color = "black"),
        axis.title.y.left = element_text(size = 12, color = "black"),
        axis.text.y.right = element_text(size = 10, color = "#4575B4"),
        axis.title.y.right= element_text(size = 11, color = "#4575B4"),
        plot.tag     = element_text(size = 12, face = "bold"))

sprintf("Figure 2C: covariation reference n = %d, lung function n = %d",
        length(c2_idx[[1]]), sum(lengths(c2_idxF)))
sprintf("series track: Spearman = %.2f (95%% CI %.2f to %.2f), P = %.4f",
        cor(fig2c_df$rel, fig2c_lf$pct, method = "spearman"),
        quantile(c2_track, 0.025), quantile(c2_track, 0.975), c2_track_p)
print(fig2c_df, row.names = FALSE, digits = 3)
print(fig2c_lf, row.names = FALSE, digits = 3)
```

### Figure 2

```{r fig_2, include=FALSE}
fig2_panel_size <- grid::unit(3.2, "in")

leg_2b <- cowplot::get_legend(panel_2b)
leg_2a <- cowplot::get_legend(panel_2a)
strip  <- function(p) egg::set_panel_size(
  p + theme(legend.position = "none", aspect.ratio = NULL),
  width = fig2_panel_size, height = fig2_panel_size)
fig2_panels <- lapply(list(panel_2a, panel_2b, panel_2c), strip)

pad_to_height <- function(g, target_h_in) {
  cur_h <- sum(grid::convertHeight(g$heights, "in", valueOnly = TRUE))
  if (cur_h < target_h_in) {
    g <- gtable::gtable_add_rows(g, grid::unit(target_h_in - cur_h, "in"))
  }
  g
}
fig2_hmax   <- max(vapply(fig2_panels, function(g)
  sum(grid::convertHeight(g$heights, "in", valueOnly = TRUE)), numeric(1)))
fig2_panels <- lapply(fig2_panels, pad_to_height, target_h_in = fig2_hmax)

top_row  <- cowplot::plot_grid(plotlist = fig2_panels, ncol = 3)
leg_row  <- cowplot::plot_grid(leg_2a, leg_2b, NULL, ncol = 3)
figure_2 <- cowplot::plot_grid(top_row, leg_row, ncol = 1, rel_heights = c(1, 0.2))
ggsave("figures/figure_2.pdf", figure_2, width = 15, height = 7,
       device = cairo_pdf)
```

### Figure 3B

```{r fig_3B_donut, include=FALSE}
panel_3b <- ggplot(inhibition_counts, aes(ymax = ymax, ymin = ymin,
                                            xmax = 4, xmin = 3,
                                            fill = inhibition)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") +
  scale_fill_manual(
    values = c("0" = "#D9D9D9", "1" = "#9ECAE1",
               "2" = "#4292C6", "3" = "#08519C"),
    name   = "ZOI Range (cm)",
    labels = c("0", "0–1", "1–2", ">2")
  ) +
  annotate("text", x = 2, y = 0, label = paste("N =", formatted_total),
           size = 3, fontface = "bold", hjust = 0.5, vjust = 0.5)
```

### Figure 3C

```{r fig_3C, include=FALSE}
cf_path_targets <- c("*A. xylosoxidans*", "*B. cenocepacia*", "*B. cepacia*",
                     "*P. aeruginosa*", "*S. aureus*")

target_label_map <- setNames(
  sapply(levels(target_hits$target), function(t) {
    clean <- gsub("^\\*|\\*$", "", t)
    if (t %in% cf_path_targets) {
      paste0("<span style='color:#E69F00; font-weight:bold; font-style:italic'>",
             clean, "</span>")
    } else {
      paste0("<i>", clean, "</i>")
    }
  }),
  levels(target_hits$target)
)

n_labels_target <- target_hits %>%
  group_by(target) %>%
  summarize(n_total = first(n_total), hit_rate = first(hit_rate), .groups = "drop")

target_totals <- target_hits %>%
  group_by(target) %>%
  summarize(total_prop = sum(prop), .groups = "drop")

panel_3c <- ggplot(target_hits, aes(x = target, y = prop,
                                     fill = factor(inhibition, levels = 0:3))) +
  gg_theme +
  geom_col(position = position_stack(reverse = TRUE)) +
  geom_col(data = target_totals, aes(x = target, y = total_prop),
           fill = NA, color = "black", linewidth = 0.2, inherit.aes = FALSE) +
  geom_text(
    data = n_labels_target,
    aes(x = target, y = hit_rate, label = paste0("N=", scales::comma(n_total))),
    inherit.aes = FALSE,
    angle = 45, hjust = 0, vjust = -0.4, size = 2.8
  ) +
  scale_fill_manual(
    values = c("0" = "#D9D9D9", "1" = "#9ECAE1", "2" = "#4292C6", "3" = "#08519C"),
    name   = "ZOI Range (cm)",
    breaks = c("0", "1", "2", "3"),
    labels = c("0 cm", "0–1 cm", "1–2 cm", ">2 cm"),
    limits = c("0", "1", "2", "3"),
    drop   = FALSE
  ) +
  scale_x_discrete(labels = target_label_map) +
  scale_y_continuous(breaks = seq(0, 60, by = 10),
                     limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.22))) +
  theme(axis.text.x     = ggtext::element_markdown(angle = 45, hjust = 1, size = 12),
        axis.text.y     = element_text(color = "black", size = 10),
        axis.title.x    = element_text(color = "black", size = 12),
        axis.title.y    = element_text(color = "black", size = 12),
        legend.position = "bottom",
        legend.text     = element_text(size = 9),
        legend.title    = element_text(size = 10)) +
  guides(fill = guide_legend(nrow = 1,
                             override.aes = list(color = "black", linewidth = 0.2))) +
  ylab("Inhibition Rate (%)") +
  xlab("Target")
```

### Figure 3

```{r fig_3, include=FALSE}
# Placeholder for 3A — same display size as 3B (donut)
panel_3a_placeholder <- ggplot() +
  theme_void() +
  theme(panel.background = element_rect(fill = "white",
                                        color = "gray80",
                                        linewidth = 0.3)) +
  annotate("text", x = 0.5, y = 0.5, label = "3A placeholder",
           size = 3, color = "gray50")

# Build the shared legend from a dummy plot so all 4 ZOI categories
leg_3_data <- data.frame(
  x = 1:4, y = 1,
  inhibition = factor(c("0", "1", "2", "3"), levels = c("0", "1", "2", "3"))
)
leg_3_dummy <- ggplot(leg_3_data, aes(x = x, y = y, fill = inhibition)) +
  geom_col(color = "black", linewidth = 0.2) +
  scale_fill_manual(
    values = c("0" = "#D9D9D9", "1" = "#9ECAE1", "2" = "#4292C6", "3" = "#08519C"),
    name   = "ZOI Range (cm)",
    breaks = c("0", "1", "2", "3"),
    labels = c("0 cm", "0\u20131 cm", "1\u20132 cm", ">2 cm"),
    drop   = FALSE
  ) +
  guides(fill = guide_legend(nrow = 1,
                             override.aes = list(color = "black", linewidth = 0.2))) +
  theme(legend.position = "bottom",
        legend.text     = element_text(size = 9),
        legend.title    = element_text(size = 10))
leg_3 <- cowplot::get_legend(leg_3_dummy)

top_row_3 <- cowplot::plot_grid(
  panel_3a_placeholder + labs(tag = "A") +
    theme(plot.tag = element_text(size = 12, face = "bold")),
  panel_3b + labs(tag = "B") +
    theme(plot.tag = element_text(size = 12, face = "bold")),
  ncol = 2
)
panel_3c_grob <- ggplotGrob(
  panel_3c + labs(tag = "C") +
    theme(plot.tag = element_text(size = 12, face = "bold"),
          legend.position = "none")
)
panel_3c_sized <- fix_panel_size_grob(panel_3c_grob,
                                      grid::unit(14.676, "cm"),
                                      grid::unit(7.897,  "cm"))

bottom_row_3 <- cowplot::plot_grid(panel_3c_sized, ncol = 1)

figure_3 <- cowplot::plot_grid(top_row_3, bottom_row_3, leg_3,
                               ncol = 1, rel_heights = c(1, 1.2, 0.12))

ggsave("figures/figure_3.pdf", figure_3, units = "in",
       width = 8, height = 9, device = cairo_pdf)
```

### Figure 4

```{r fig_4_xstreak, include=FALSE}
SY_MIN_ASSAYS <- 10

sy_genera <- unique(c(pairwise_results$genus1, pairwise_results$genus2))

sy_assays <- bioassay_data %>%
  left_join(cf_isolates %>% dplyr::select(isolate, genus), by = "isolate") %>%
  dplyr::rename(producer = genus) %>%
  mutate(target_genus = sub(" .*$", "", target)) %>%
  filter(!is.na(producer), producer %in% sy_genera,
         target_genus %in% sy_genera, producer != target_genus)

sy_dir <- sy_assays %>%
  group_by(producer, target_genus) %>%
  summarize(assays = n(), pct_inh = 100 * mean(inhibition >= 1), .groups = "drop") %>%
  filter(assays >= SY_MIN_ASSAYS)

sy_key <- function(a, b) paste(pmin(a, b), pmax(a, b), sep = "|")
sy_pairs <- sy_dir %>%
  mutate(pair = sy_key(producer, target_genus)) %>%
  group_by(pair) %>%
  slice_max(pct_inh, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(bx = pairwise_results$bx[match(pair,
           sy_key(pairwise_results$genus1, pairwise_results$genus2))]) %>%
  filter(!is.na(bx))

SY_BAR <- "⟞"
sy_lv  <- c(paste("Pathogen",      SY_BAR, "Pathogen"),
            paste("Pathogen",      SY_BAR, "Oropharyngeal"),
            paste("Oropharyngeal", SY_BAR, "Pathogen"),
            paste("Oropharyngeal", SY_BAR, "Oropharyngeal"))
sy_pp <- sy_pairs$producer     %in% cf_pathogens
sy_tp <- sy_pairs$target_genus %in% cf_pathogens
sy_pairs$pair_type <- factor(
  ifelse(sy_pp &  sy_tp, sy_lv[1],
  ifelse(sy_pp & !sy_tp, sy_lv[2],
  ifelse(!sy_pp & sy_tp, sy_lv[3], sy_lv[4]))), levels = sy_lv)

sy_ct  <- suppressWarnings(
  cor.test(sy_pairs$pct_inh, sy_pairs$bx, method = "spearman"))
sy_lab <- paste0("italic(r)[s]== ", sprintf("%.2f", sy_ct$estimate),
                 "~~italic(P)== ", sprintf("%.3f", sy_ct$p.value))

sy_cols <- setNames(c("#E69F00", "#4575B4", "#56B4E9", "gray55"), sy_lv)

figure_4 <- ggplot(sy_pairs, aes(pct_inh, bx)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "gray45", linewidth = 0.45) +
  geom_smooth(method = "lm", formula = y ~ x, color = "#D73027",
              fill = "#D73027", alpha = 0.16, linewidth = 1.0) +
  geom_point(aes(color = pair_type, size = assays), alpha = 0.9) +
  scale_color_manual(values = sy_cols) +
  scale_size_continuous(range = c(1.4, 4.6), breaks = c(20, 100, 500, 1200)) +
  scale_x_continuous("Target Inhibition in Cross-Streak Assays (%)",
                     breaks = seq(0, 80, 20)) +
  scale_y_continuous("Covariation Coefficient",
                     breaks = seq(-0.1, 0.5, 0.1),
                     labels = sprintf("%.1f", seq(-0.1, 0.5, 0.1))) +
  coord_cartesian(xlim = c(-6, 92), ylim = c(-0.18, 0.60)) +
  theme_classic(base_size = 11) +
  theme(aspect.ratio  = 1,
        axis.line     = element_blank(),
        panel.border  = element_rect(color = "black", fill = NA, linewidth = 0.35),
        axis.text     = element_text(size = 9.5, color = "black"),
        axis.title    = element_text(size = 11, color = "black"),
        axis.ticks    = element_line(color = "black", linewidth = 0.3),
        legend.position   = "bottom",
        legend.title      = element_text(size = 9.2),
        legend.text       = element_text(size = 9.2),
        legend.key        = element_blank(),
        legend.background = element_blank(),
        legend.key.height = unit(0.32, "cm"),
        legend.box        = "vertical",
        legend.spacing.y  = unit(0.5, "cm"),
        legend.margin     = margin(t = 0, b = 0)) +
  guides(color = guide_legend(title = NULL, nrow = 2, order = 1,
                              override.aes = list(size = 2.4)),
         size  = guide_legend(title = "Number of Assays", nrow = 1, order = 2,
                              title.position = "left", title.vjust = 0.5))

SY_SYMFONT <- "Segoe UI Symbol"
SY_GAP     <- 4.5   # pt of space on each side of the inhibition bar
SY_RAISE   <- 0.6   # pt lift so the bar centres on the cap height, not the baseline
sy_metrics <- function(s, fam, fs, K = 100) {
  m <- systemfonts::shape_string(s, family = fam, size = fs * K)$metrics
  list(adv = m$width / K, lb = m$left_bearing / K, rb = m$right_bearing / K)
}
sy_mixed <- function(g0) {
  parts <- trimws(strsplit(g0$label, SY_BAR, fixed = TRUE)[[1]])
  fs  <- if (is.null(g0$gp$fontsize)) 9.2 else g0$gp$fontsize
  fam <- if (is.null(g0$gp$fontfamily) || !nzchar(g0$gp$fontfamily)) "Arial"
         else g0$gp$fontfamily
  m1 <- sy_metrics(parts[1], fam, fs)
  mb <- sy_metrics(SY_BAR, SY_SYMFONT, fs)
  m2 <- sy_metrics(parts[2], fam, fs)
  # place on ink edges, not advance widths, so the gaps either side of the bar
  # are equal to the eye
  xb  <- (m1$adv - m1$rb) + SY_GAP - mb$lb
  x2  <- (xb + mb$adv - mb$rb) + SY_GAP - m2$lb
  off <- c(0, xb, x2)
  txt <- c(parts[1], SY_BAR, parts[2])
  gps <- lapply(c("", SY_SYMFONT, ""), function(f) {
    g <- g0$gp; if (nzchar(f)) g$fontfamily <- f; g
  })
  gt <- grid::gTree(children = do.call(grid::gList, mapply(function(s, g, o, r)
    grid::textGrob(s, x = g0$x + grid::unit(o, "pt"),
                   y = g0$y + grid::unit(r, "pt"),
                   hjust = 0, vjust = g0$vjust, gp = g),
    txt, gps, off, c(0, SY_RAISE, 0), SIMPLIFY = FALSE)))
  attr(gt, "sy_w") <- x2 + m2$adv
  gt
}
sy_fix <- function(x) {
  if (inherits(x, "gtable")) {
    for (k in seq_along(x$grobs)) {
      x$grobs[[k]] <- sy_fix(x$grobs[[k]])
      w  <- attr(x$grobs[[k]], "sy_w")
      cl <- x$layout$l[k]
      if (!is.null(w)) {
        x$widths[cl] <- grid::unit.pmax(x$widths[cl], grid::unit(w + 5.5, "pt"))
      }
    }
    x
  }
  else if (inherits(x, "gTree")) {
    if (!is.null(x$children)) {
      x$children <- do.call(grid::gList, lapply(x$children, sy_fix))
      w <- vapply(x$children, function(ch) {
        a <- attr(ch, "sy_w"); if (is.null(a)) NA_real_ else a
      }, numeric(1))
      if (any(!is.na(w))) attr(x, "sy_w") <- max(w, na.rm = TRUE)
    }
    x
  } else if (inherits(x, "text") && length(x$label) == 1 &&
             grepl(SY_BAR, x$label, fixed = TRUE)) sy_mixed(x) else x
}

dir.create("figures", showWarnings = FALSE)
figure_4_sized <- egg::set_panel_size(figure_4 + theme(aspect.ratio = NULL),
                                      width  = fig2_panel_size,
                                      height = fig2_panel_size)
figure_4_sized <- sy_fix(figure_4_sized)
ggsave("figures/figure_4.pdf", figure_4_sized,
       width  = sum(grid::convertWidth(figure_4_sized$widths, "in", valueOnly = TRUE)),
       height = sum(grid::convertHeight(figure_4_sized$heights, "in", valueOnly = TRUE)),
       device = cairo_pdf)

sprintf("Figure 4: %d genus pairs, Spearman rho = %.3f, P = %.4f",
        nrow(sy_pairs), sy_ct$estimate, sy_ct$p.value)
print(as.data.frame(table(sy_pairs$pair_type)))
```

### Figure 5A

```{r fig_5A, include=FALSE}
fig5_lead   <- c("*Pseudomonas*", "*Stenotrophomonas*", "*Escherichia*", "*Staphylococcus*")
fig5_others <- setdiff(unique(as.character(isolate_inhibition$genus)), fig5_lead)
fig5_order  <- c(fig5_lead, fig5_others)

isolate_inhibition_plot <- isolate_inhibition %>%
  mutate(genus = factor(as.character(genus), levels = fig5_order))

isolate_inhibition_figure <- ggplot(isolate_inhibition_plot, aes(x = genus, y = percent, fill = inhibition)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  geom_col(
    data = isolate_inhibition_plot %>%
      group_by(genus) %>% summarize(total_pct = sum(percent), .groups = "drop"),
    aes(x = genus, y = total_pct),
    fill = NA, color = "black", linewidth = 0.2, inherit.aes = FALSE) +
  gg_theme +
  scale_fill_manual(
    values = c("0" = "#D9D9D9", "1" = "#9ECAE1", "2" = "#4292C6", "3" = "#08519C"),
    name   = "ZOI Range (cm)",
    breaks = c("0", "1", "2", "3"),
    labels = c("0 cm", "0-1 cm", "1-2 cm", ">2 cm"),
    limits = c("0", "1", "2", "3"),
    drop   = FALSE
  ) +
  scale_x_discrete(labels = function(x) {
    m <- isolate_inhibition_plot %>% distinct(genus, N_isolates) %>%
      mutate(
        is_path = sapply(as.character(genus), function(g) any(sapply(cf_pathogens, grepl, x = g))),
        label   = ifelse(
          is_path,
          paste0("<span style='color:#E69F00'>", genus, "<br/>(n=", N_isolates, ")</span>"),
          paste0(genus, "<br/>(n=", N_isolates, ")")
        )
      )
    setNames(m$label, m$genus)[x]
  }) +
  geom_text(
    data = isolate_inhibition_plot %>% group_by(genus) %>% summarize(N_interactions = sum(count), .groups = "drop"),
    aes(x = genus, y = 1.06, label = paste0("N=", scales::comma(N_interactions))),
    inherit.aes = FALSE,
    angle = 45,
    vjust = -0.4,
    size  = 2.8
  ) +
  scale_y_continuous(
    breaks  = seq(0, 1, by = 0.2),
    labels  = scales::number_format(accuracy = 1, scale = 100),
    limits  = c(0, 1.18),
    expand  = expansion(mult = c(0, 0.02))
  ) +
  xlab(NULL) +
  ylab("ZOI Category (%)") +
  theme(
    axis.text.x  = ggtext::element_markdown(angle = 45, hjust = 1, size = 12, lineheight = 1.0),
    axis.text.y  = element_text(color = "black", size = 10),
    axis.title.x = element_text(color = "black", size = 12),
    axis.title.y = element_text(color = "black", size = 12),
    legend.position = "none"
  )

leg_5a <- cowplot::get_legend(
  isolate_inhibition_figure + theme(legend.position = "bottom",
                                     legend.text  = element_text(size = 9),
                                     legend.title = element_text(size = 10))
)
```

### Figure 5B

```{r fig_5B, include=FALSE}
total_inhibition_per_isolate <- total_inhibition_per_isolate %>%
  mutate(genus = factor(as.character(genus), levels = fig5_order))

chi_pairs_5b <- combn(fig5_order, 2, simplify = FALSE)
chi_pairs_5b_df <- do.call(rbind, lapply(chi_pairs_5b, function(g) {
  d <- isolate_inhibition_for_chi %>%
    filter(genus %in% g) %>%
    mutate(genus = droplevels(factor(genus)))
  tab <- xtabs(count ~ genus + inhibition, data = d)
  data.frame(g1 = g[1], g2 = g[2],
             p = suppressWarnings(
               chisq.test(tab, simulate.p.value = TRUE, B = 10000)$p.value))
}))
chi_pairs_5b_df$padj <- p.adjust(chi_pairs_5b_df$p, method = "BH")

chi_padj_vec <- chi_pairs_5b_df$padj
names(chi_padj_vec) <- paste(chi_pairs_5b_df$g1, chi_pairs_5b_df$g2, sep = "-")
chi_letters_obj <- multcompLetters(chi_padj_vec, threshold = 0.05)

raw_chi    <- chi_letters_obj$Letters[fig5_order]
chars_chi  <- unique(unlist(strsplit(raw_chi, "")))
lmap_chi   <- setNames(letters[seq_along(chars_chi)], chars_chi)
inh_letters <- sapply(raw_chi, function(l)
  paste(sort(lmap_chi[strsplit(l, "")[[1]]]), collapse = ""))

raw_fl     <- setNames(fligner_letters_df$letters, fligner_letters_df$genus)[fig5_order]
chars_fl   <- unique(unlist(strsplit(raw_fl, "")))
lmap_fl    <- setNames(letters[seq_along(chars_fl)], chars_fl)
disp_letters <- sapply(raw_fl, function(l)
  paste(sort(lmap_fl[strsplit(l, "")[[1]]]), collapse = ""))

kw_letters_df <- data.frame(
  genus   = fig5_order,
  letters = paste0(inh_letters, "/", disp_letters),
  row.names = NULL
)

n_lab <- total_inhibition_per_isolate %>%
  count(genus, name = "n") %>%
  mutate(
    is_path = sapply(as.character(genus), function(g) any(sapply(cf_pathogens, grepl, x = g))),
    label   = ifelse(
      is_path,
      paste0("<span style='color:#E69F00'>", genus, "<br/>(n=", n, ")</span>"),
      paste0(genus, "<br/>(n=", n, ")")
    )
  )

lab_map <- setNames(n_lab$label, n_lab$genus)

pos <- aggregate(z_overall ~ genus, data = total_inhibition_per_isolate, FUN = max)
pos$y <- pos$z_overall + 0.30

ann <- merge(pos[, c("genus","y")], kw_letters_df, by = "genus", all.x = TRUE)

genus_lvls <- levels(total_inhibition_per_isolate$genus)
genus_cols <- setNames(
  vapply(genus_lvls, function(g) {
    if (any(vapply(cf_pathogens, grepl, logical(1), x = g))) "#E69F00" else "gray60"
  }, character(1)),
  genus_lvls
)

z_outliers <- total_inhibition_per_isolate %>%
  group_by(genus) %>%
  mutate(.q1 = quantile(z_overall, 0.25, na.rm = TRUE),
         .q3 = quantile(z_overall, 0.75, na.rm = TRUE),
         .iqr = .q3 - .q1) %>%
  filter(z_overall < .q1 - 1.5 * .iqr | z_overall > .q3 + 1.5 * .iqr) %>%
  ungroup()

z_score_per_isolate_plot <- ggplot(data = total_inhibition_per_isolate,
                                    aes(x = genus, y = z_overall, fill = genus)) +
  gg_theme +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_boxplot(outlier.shape = NA, color = "black", linewidth = 0.3, alpha = 0.7) +
  geom_point(data = z_outliers, aes(fill = genus),
             shape = 21, color = "black", stroke = 0.2,
             size = 1.2, alpha = 0.85,
             position = ggbeeswarm::position_quasirandom(width = 0.25)) +
  scale_fill_manual(values = genus_cols) +
  scale_x_discrete(labels = lab_map) +
  xlab("Isolate Genus") +
  ylab("Inhibition Z-Score") +
  theme(
    axis.text.x  = ggtext::element_markdown(angle = 45, hjust = 1, size = 12, lineheight = 1.0),
    axis.text.y  = element_text(color = "black", size = 10),
    axis.title.x = element_text(color = "black", size = 12),
    axis.title.y = element_text(color = "black", size = 12),
    legend.position = "none"
  ) +
  geom_text(
    data = ann,
    aes(x = genus, y = y, label = letters),
    inherit.aes = FALSE,
    size = 3.5
  ) +
  expand_limits(y = max(ann$y, na.rm = TRUE) + 0.2)

#ggsave("./figures/z_score_per_isolate_plot.pdf", z_score_per_isolate_plot, units = "in", height = 4, width = 6.47)
```

### Figure 5

```{r fig_5, include=FALSE}
aligned_5 <- cowplot::align_plots(isolate_inhibition_figure,
                                   z_score_per_isolate_plot,
                                   align = "v", axis = "lr")

combined_inhibition_figure <- ggdraw(
  plot_grid(aligned_5[[1]], leg_5a, aligned_5[[2]],
            ncol = 1, rel_heights = c(1, 0.1, 1))
) +
  draw_label("A", x = 0.01, y = 0.99, hjust = 0, vjust = 1, fontface = "bold", size = 14) +
  draw_label("B", x = 0.01, y = 0.475, hjust = 0, vjust = 1, fontface = "bold", size = 14)

ggsave("./figures/figure_5.pdf", combined_inhibition_figure, units = "in",
       height = 8.5, width = 6.47)
```

### Figure 7

```{r fig_7, include=FALSE}
lcms_mcol <- grep("\\.mean$", colnames(lcms_diff), value = TRUE)
lcms_M    <- as.matrix(lcms_diff[, lcms_mcol])
lcms_gmax <- apply(lcms_M, 1, max)
lcms_gmin <- apply(lcms_M, 1, function(x) { x <- x[x > 0]; if (length(x)) min(x) else NA })
lcms_feat <- lcms_diff %>%
  transmute(rt_min = rtmed, mz = mzmed, pvalue = pvalue,
            log2fc = log2(lcms_gmax / lcms_gmin), neglog10p = -log10(pvalue)) %>%
  filter(pvalue <= lcms_p_cutoff, is.finite(log2fc)) %>%
  arrange(neglog10p)

lcms_tinh <- lcms_meta$inhibition[match(lcms_ciso, lcms_meta$isolate)]
lcms_bidx <- which(lcms_ciso == "BHI"); lcms_keep <- which(lcms_ciso != "BHI")
lcms_rtall <- range(unlist(lapply(lcms_tics, function(m) range(m[, 1]))))
lcms_grid  <- seq(lcms_rtall[1], lcms_rtall[2], length.out = 3000)
lcms_blank <- rowMeans(sapply(lcms_bidx, function(i) {
  m <- lcms_tics[[i]]; approx(m[, 1], m[, 2], xout = lcms_grid, rule = 2)$y }))
lcms_bfun  <- approxfun(lcms_grid, lcms_blank, rule = 2)
lcms_mztop <- max(lcms_feat$mz)
lcms_ticdf <- bind_rows(lapply(lcms_keep, function(i) {
  m <- lcms_tics[[i]]
  data.frame(sample = i, inhibition = lcms_tinh[i], rt_min = m[, 1] / 60,
             tic = pmax(m[, 2] - lcms_bfun(m[, 1]), 0)) }))
lcms_ticdf$mz <- lcms_ticdf$tic / max(lcms_ticdf$tic) * lcms_mztop * 0.95

panel_7A <- ggplot() +
  geom_line(data = lcms_ticdf, aes(rt_min, mz, group = sample, color = inhibition),
            linewidth = 0.15, alpha = 0.5) +
  geom_point(data = lcms_feat, aes(rt_min, mz, size = log2fc),
             color = "#B00000", fill = "#B00000", shape = 21, stroke = 0.15, alpha = 0.5) +
  scale_color_manual(values = lcms_inh_pal, breaks = c("Yes", "No"), name = "Inhibition Z-Score ≥ 0") +
  scale_radius(range = c(0.2, 4), limits = c(0, 15), breaks = c(0, 5, 10, 15),
               name = expression(log[2]~"Fold Change")) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.03))) +
  labs(x = "Retention Time (min)", y = expression(italic("m/z"))) +
  fig1_theme +
  theme(legend.title = element_text(size = 10), legend.text = element_text(size = 9),
        legend.key = element_blank())

lcms_fig_w <- 7; lcms_fig_h <- 8.5; lcms_rel_h <- c(1, 1.9); lcms_rel_w <- c(1, 0.22)
lcms_plots_w <- lcms_fig_w * lcms_rel_w[1] / sum(lcms_rel_w)

lcms_node_pos <- function(hc) {
  m <- hc$merge; ord <- hc$order; cx <- numeric(nrow(m))
  for (i in seq_len(nrow(m))) {
    a <- m[i, 1]; b <- m[i, 2]
    xa <- if (a < 0) which(ord == -a) - 0.5 else cx[a]
    xb <- if (b < 0) which(ord == -b) - 0.5 else cx[b]
    cx[i] <- (xa + xb) / 2
  }
  data.frame(x = cx, y = hc$height)
}
dir.create("r_objects", showWarnings = FALSE)
if (file.exists("r_objects/lcms_pvclust.rds")) {
  lcms_pv <- readRDS("r_objects/lcms_pvclust.rds")
} else {
  set.seed(8675309)   
  lcms_pv <- pvclust(lcms_matz, method.dist = "euclidean", method.hclust = "complete",
                     nboot = lcms_n_boot, quiet = TRUE)
  saveRDS(lcms_pv, "r_objects/lcms_pvclust.rds")
}

pdf(NULL, width = lcms_plots_w, height = 3)
lcms_gA <- ggplotGrob(panel_7A + theme(legend.position = "none"))
lcms_pl <- subset(lcms_gA$layout, name == "panel")
lcms_left_mm <- convertWidth(sum(lcms_gA$widths[seq_len(min(lcms_pl$l) - 1)]), "mm", valueOnly = TRUE)
invisible(dev.off())

lcms_topann <- HeatmapAnnotation(
  HID = lcms_iso_meta$host, Inhibition = lcms_iso_meta$inhibition,
  col = list(HID = lcms_host_pal, Inhibition = lcms_inh_pal),
  simple_anno_size = unit(6, "mm"), show_annotation_name = FALSE,
  show_legend = FALSE, gap = unit(1.5, "mm"))
lcms_ht <- Heatmap(
  lcms_matz, name = "Feature Z-Score", top_annotation = lcms_topann,
  cluster_rows = TRUE, show_row_dend = FALSE, show_row_names = FALSE,
  cluster_columns = as.dendrogram(lcms_pv$hclust), column_dend_height = unit(16, "mm"),
  column_names_gp = gpar(fontsize = 10), column_names_rot = 45, col = lcms_val_col,
  use_raster = TRUE, raster_quality = 3,
  heatmap_legend_param = list(direction = "horizontal"), border = TRUE)
panel_7B <- grid.grabExpr({
  draw(lcms_ht, heatmap_legend_side = "bottom", show_annotation_legend = FALSE,
       padding = unit(c(2, lcms_left_mm, 2, 19), "mm"))
  decorate_column_dend("Feature Z-Score", {
    np <- lcms_node_pos(lcms_pv$hclust)
    grid.text(round(lcms_pv$edges$au * 100), x = unit(np$x, "native"), y = unit(np$y, "native"),
              just = c("center", "bottom"), gp = gpar(fontsize = 5.5, col = "black"))
  })
  decorate_annotation("HID",
    grid.text("HID", x = unit(1, "npc") + unit(1, "mm"), just = "left", gp = gpar(fontsize = 10, fontface = "bold")))
  decorate_annotation("Inhibition",
    grid.text("Inhibition", x = unit(1, "npc") + unit(1, "mm"), just = "left", gp = gpar(fontsize = 10, fontface = "bold")))
}, width = lcms_plots_w, height = lcms_fig_h * lcms_rel_h[2] / sum(lcms_rel_h))

lcms_legA <- get_legend(panel_7A + guides(color = "none") +
  theme(legend.position = "right", legend.box = "vertical",
        legend.title = element_text(size = 10), legend.text = element_text(size = 9)))
lcms_legB <- packLegend(
  Legend(title = "HID", labels = names(lcms_host_pal),
         legend_gp = gpar(fill = unname(lcms_host_pal)),
         title_gp = gpar(fontsize = 10, fontface = "bold"), labels_gp = gpar(fontsize = 9)),
  Legend(title = "Inhibition\nZ-Score ≥ 0", labels = c("Yes", "No"),
         legend_gp = gpar(fill = unname(lcms_inh_pal[c("Yes", "No")])),
         title_gp = gpar(fontsize = 10, fontface = "bold"), labels_gp = gpar(fontsize = 9)),
  gap = unit(5, "mm"))
lcms_legB_grob <- grid.grabExpr(draw(lcms_legB))

lcms_plots_col <- plot_grid(panel_7A + theme(legend.position = "none",
                              plot.margin = margin(5.5, 53.9, 5.5, 5.5, "pt")), panel_7B,
                            ncol = 1, labels = c("A", "B"), label_size = 12,
                            label_fontface = "bold", rel_heights = lcms_rel_h)
lcms_legs_col  <- plot_grid(lcms_legA, lcms_legB_grob, ncol = 1, rel_heights = lcms_rel_h)
figure_7       <- plot_grid(lcms_plots_col, lcms_legs_col, ncol = 2, rel_widths = lcms_rel_w)

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_7.pdf", figure_7, width = lcms_fig_w, height = lcms_fig_h, device = cairo_pdf)
```

### Figure 8

```{r fig_8, include=FALSE}
local({
  PD  <- file.path("derivedData", "genome_comparison", "plotdata")
  FIG <- "figures"
  dir.create(FIG, showWarnings = FALSE)

  meta   <- read.csv(file.path(PD, "meta.csv"), stringsAsFactors = FALSE)
  legend <- read.csv(file.path(PD, "bgc_legend.csv"), stringsAsFactors = FALSE)

  # palette follows this document (RdYlBu / Okabe-Ito)
  VAR_CLASSES <- c("Frameshift", "Missense", "Silent", "Intergenic")
  VAR_COL <- c("Frameshift" = "#D73027", "Missense" = "#E69F00",
               "Silent" = "#999999", "Intergenic" = "#CC79A7")

  # radii (unit circle)
  R_BGC_OUT <- 1.00; R_BGC_IN <- 0.90
  R_MUT_OUT <- 0.86; R_MUT_IN <- 0.72
  R_NUM     <- 1.13
  R_MB      <- 1.21

  xy <- function(r, f) list(x = r * sin(2 * pi * f), y = r * cos(2 * pi * f))
  tf <- function(f, o, off) ((o * f) + off) %% 1       # reorient panel B onto A

  arc_poly <- function(f1, f2, ri, ro, col) {
    if (f2 < f1) f2 <- f2 + 1
    ff  <- seq(f1, f2, length.out = max(2, round((f2 - f1) * 400)))
    ang <- 2 * pi * ff
    polygon(c(ro * sin(ang), rev(ri * sin(ang))),
            c(ro * cos(ang), rev(ri * cos(ang))), col = col, border = NA)
  }

  spread_ang <- function(ang, mingap) {                # de-overlap numeric labels
    ord <- order(ang); a <- ang[ord]; n <- length(a)
    for (it in 1:400) {
      moved <- FALSE
      for (i in 1:n) {
        j <- if (i == n) 1 else i + 1
        gap <- (a[j] - a[i]) %% (2 * pi)
        if (gap < mingap) {
          d <- (mingap - gap) / 2
          a[i] <- (a[i] - d) %% (2 * pi); a[j] <- (a[j] + d) %% (2 * pi); moved <- TRUE
        }
      }
      if (!moved) break
    }
    out <- numeric(n); out[ord] <- a; out
  }

  draw_panel <- function(donor, panel) {
    m   <- meta[meta$donor == donor, ]
    o <- m$orient; off <- m$offset
    bgc <- read.csv(file.path(PD, sprintf("bgc_%s.csv", donor)), stringsAsFactors = FALSE)
    mut <- read.csv(file.path(PD, sprintf("mut_%s.csv", donor)), stringsAsFactors = FALSE)
    mut <- mut[mut$noninh_specific == 1, , drop = FALSE]          # only variants in the non-inhibitory genome
    L   <- m$length

    par(mar = c(0.4, 0.4, 0.4, 0.4))
    plot(NA, xlim = c(-1.32, 1.32), ylim = c(-1.30, 1.34), asp = 1,
         axes = FALSE, xlab = "", ylab = "")

    arc_poly(0, 1, R_BGC_IN, R_BGC_OUT, col = "#ECEAE2")           # backbone

    for (mb in 0:floor(L / 1e6)) {                                 # Mb ticks
      f <- mb * 1e6 / L
      p1 <- xy(R_BGC_OUT, f); p2 <- xy(R_BGC_OUT + 0.035, f)
      segments(p1$x, p1$y, p2$x, p2$y, col = "#9A988F", lwd = 0.9)
      pl <- xy(R_MB, f); text(pl$x, pl$y, mb, col = "#6b6a64", cex = 0.95)
    }

    for (i in seq_len(nrow(bgc))) {                                # BGC arcs
      a <- tf(bgc$f_start[i], o, off); b <- tf(bgc$f_end[i], o, off)
      if (o == -1) { tmp <- a; a <- b; b <- tmp }
      arc_poly(a, b, R_BGC_IN, R_BGC_OUT, col = bgc$color[i])
    }

    ang0 <- 2 * pi * tf(bgc$f_mid, o, off)
    angL <- spread_ang(ang0, mingap = 0.115)
    for (i in seq_len(nrow(bgc))) {
      a0 <- xy(R_BGC_OUT, ang0[i] / (2 * pi)); aL <- xy(R_NUM, angL[i] / (2 * pi))
      segments(a0$x, a0$y, aL$x, aL$y, col = "#B4B2A9", lwd = 0.7)
      text(aL$x, aL$y, bgc$num[i], cex = 1.15, col = "#2C2C2A")
    }

    for (i in seq_len(nrow(mut))) {
      f <- tf(mut$f[i], o, off)
      p1 <- xy(R_MUT_IN, f); p2 <- xy(R_MUT_OUT, f)
      segments(p1$x, p1$y, p2$x, p2$y, col = VAR_COL[[mut$class[i]]], lwd = 2.0)
      if (mut$inbgc[i] == 1) {                                    # BGC-associated
        ps <- xy(R_MUT_IN - 0.035, f)
        text(ps$x, ps$y, "*", col = "#2C2C2A", cex = 1.8, font = 2)
      }
    }

    text(0,  0.15, m$reference, cex = 2.0, font = 2)              
    text(0,  0.01, donor, cex = 1.15, col = "#5F5E5A")
    text(0, -0.12, sprintf("%.2f Mb", L / 1e6), cex = 1.1, col = "#5F5E5A")
    text(0, -0.25, sprintf("%d variants", nrow(mut)), cex = 1.1, col = "#5F5E5A")
    text(-1.30, 1.30, panel, cex = 2.0, font = 2, adj = 0)
  }

  draw_mut_key <- function() {
    par(mar = c(0, 0.5, 0, 0.5))
    plot(NA, xlim = c(0, 1), ylim = c(0, 1), axes = FALSE, xlab = "", ylab = "")
    cx <- 1.1; segw <- 0.014; gap <- 0.03               
    x <- 0.01
    for (j in seq_along(VAR_CLASSES)) {                 
      segments(x, 0.5, x + segw, 0.5, col = VAR_COL[[VAR_CLASSES[j]]], lwd = 5)
      text(x + segw + 0.006, 0.5, VAR_CLASSES[j], adj = 0, cex = cx, col = "#000000")
      x <- x + segw + 0.006 + strwidth(VAR_CLASSES[j], cex = cx) + gap
    }
    text(x, 0.5, "*", adj = 0, cex = cx * 1.3, font = 2, col = "#2C2C2A")  # BGC glyph
    text(x + strwidth("* ", cex = cx), 0.5, "BGC-Associated", adj = 0, cex = cx, col = "#000000")
  }

  draw_bgc_legend <- function() {                      
    par(mar = c(0.1, 0.5, 0.1, 0.5))
    plot(NA, xlim = c(0, 1), ylim = c(0, 1), axes = FALSE, xlab = "", ylab = "")
    n <- nrow(legend); ncols <- 9
    text(0.0, 0.93, "BGCs", adj = 0, cex = 1.2, font = 2, col = "#000000")
    colw <- 1 / ncols; swx <- 0.008; swy <- 0.06; y_row <- c(0.58, 0.20)
    for (i in seq_len(n)) {
      r  <- if (i <= ncols) 1 else 2
      cc <- if (i <= ncols) i else i - ncols
      x <- (cc - 1) * colw + 0.004; y <- y_row[r]
      rect(x, y - swy, x + swx, y + swy, col = legend$color[i], border = "#00000030", lwd = 0.4)
      text(x + swx + 0.006, y, sprintf("%d. %s", legend$num[i], legend$label[i]),
           adj = 0, cex = 0.82, col = "#000000")
    }
  }

  render <- function(open_dev, ext) {
    open_dev()
    layout(matrix(c(1, 2,
                    3, 3,
                    4, 4), nrow = 3, byrow = TRUE),
           widths = c(1, 1), heights = c(1, 0.12, 0.22))
    draw_panel("HID07", "A")
    draw_panel("HID08", "B")
    draw_mut_key()               
    draw_bgc_legend()            
    dev.off()
  }

  render(function() cairo_pdf(file.path(FIG, "figure_8.pdf"), width = 15, height = 8.7,
                              pointsize = 12), "pdf")
})
```

### Tables S6 and S7

```{r tables_S6_S7, include=FALSE}
lcms_gc_tab <- file.path("derivedData", "genome_comparison", "tables")
dir.create("tables", showWarnings = FALSE)

build_variant_table <- function(donor, outfile) {
  v <- read.delim(file.path(lcms_gc_tab, paste0(donor, "_differences.tsv")),
                  check.names = FALSE, stringsAsFactors = FALSE, colClasses = "character")
  b <- read.delim(file.path(lcms_gc_tab, paste0(donor, "_variants_in_BGCs.tsv")),
                  check.names = FALSE, stringsAsFactors = FALSE)
  v$TYPE <- c(snp = "SNP", ins = "Insertion", del = "Deletion")[v$TYPE]
  v$BGC_associated <- ifelse(as.integer(v$POS) %in% b$position, "Yes", "No")
  v$BGC <- b$BGC[match(as.integer(v$POS), b$position)]; v$BGC[is.na(v$BGC)] <- ""
  ren <- c(CHROM = "Contig", POS = "Position", TYPE = "Type", REF = "Ref", ALT = "Alt",
           EFFECT = "Effect", LOCUS_TAG = "Locus_tag", GENE = "Gene", PRODUCT = "Product")
  hit <- names(v) %in% names(ren); names(v)[hit] <- ren[names(v)[hit]]
  v[is.na(v)] <- ""                                  
  write.csv(v, outfile, row.names = FALSE)
  v
}
table_S6 <- build_variant_table("HID07", file.path("tables", "Table_S6.csv"))  # HID07
table_S7 <- build_variant_table("HID08", file.path("tables", "Table_S7.csv"))  # HID08
```

### Figure S1A

```{r fig_S1A, include=FALSE}
pathogen_dominant_obs <- per_sample_ra_lab %>%
  filter(ra > 50, group == "pathogen") %>%
  mutate(taxon = factor(as.character(taxon), levels = pathogen_in_table))

pw_pathogen <- pairwise.wilcox.test(pathogen_dominant_obs$ra,
                                     pathogen_dominant_obs$taxon,
                                     p.adjust.method = "BH", exact = FALSE)
pw_pathogen_mat <- pw_pathogen$p.value
pw_pathogen_idx <- which(!is.na(pw_pathogen_mat), arr.ind = TRUE)
pw_pathogen_p   <- pw_pathogen_mat[pw_pathogen_idx]
names(pw_pathogen_p) <- paste(rownames(pw_pathogen_mat)[pw_pathogen_idx[, "row"]],
                              colnames(pw_pathogen_mat)[pw_pathogen_idx[, "col"]],
                              sep = "-")
pathogen_letters <- multcompLetters(pw_pathogen_p, threshold = 0.05)

raw_pathogen_letters <- pathogen_letters$Letters[pathogen_in_table]
pathogen_chars       <- unique(unlist(strsplit(raw_pathogen_letters, "")))
pathogen_letter_map  <- setNames(letters[seq_along(pathogen_chars)], pathogen_chars)
pathogen_letters_df  <- data.frame(
  taxon  = pathogen_in_table,
  letter = sapply(raw_pathogen_letters, function(l)
    paste(sort(pathogen_letter_map[strsplit(l, "")[[1]]]), collapse = "")),
  row.names = NULL
)

pathogen_letter_pos <- pathogen_dominant_obs %>%
  group_by(taxon) %>%
  summarize(y_pos = max(ra) + 2, .groups = "drop") %>%
  left_join(pathogen_letters_df, by = "taxon")

n_per_pathogen <- pathogen_dominant_obs %>%
  count(taxon) %>%
  mutate(taxon = as.character(taxon))
n_pathogen_map <- setNames(n_per_pathogen$n, n_per_pathogen$taxon)

pathogen_axis_labels <- setNames(
  vapply(pathogen_in_table, function(g) {
    n <- n_pathogen_map[[g]]
    if (is.null(n) || is.na(n)) n <- 0
    paste0("<span style='color:#E69F00;'>",
           "<span style='font-weight:bold; font-style:italic'>", g, "</span>",
           "<br/>(n=", scales::comma(n), ")",
           "</span>")
  }, character(1)),
  pathogen_in_table
)

panel_S1a_outliers <- pathogen_dominant_obs %>%
  group_by(taxon) %>%
  mutate(.q1 = quantile(ra, 0.25, na.rm = TRUE),
         .q3 = quantile(ra, 0.75, na.rm = TRUE),
         .iqr = .q3 - .q1) %>%
  filter(ra < .q1 - 1.5 * .iqr | ra > .q3 + 1.5 * .iqr) %>%
  ungroup()

panel_S1a <- ggplot(pathogen_dominant_obs, aes(x = taxon, y = ra, fill = taxon)) +
  geom_boxplot(alpha = 0.7, width = 0.5, notch = TRUE, outlier.shape = NA) +
  geom_point(data = panel_S1a_outliers, aes(fill = taxon),
             shape = 21, color = "black", stroke = 0.2,
             size = 1, alpha = 0.85) +
  geom_text(data = pathogen_letter_pos,
            aes(x = taxon, y = y_pos, label = letter),
            inherit.aes = FALSE, size = 3) +
  scale_fill_manual(values = pathogen_palette, guide = "none") +
  scale_x_discrete(labels = pathogen_axis_labels) +
  scale_y_continuous(limits = c(45, 110), breaks = seq(50, 100, by = 10),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "Dominant Pathogen", y = "Relative Abundance (%)") +
  fig1_theme +
  theme(axis.text.x = ggtext::element_markdown(angle = 45, hjust = 1, size = 10))
```

### Figure S1B

```{r fig_S1B, include=FALSE}
group_cols <- c("Pathogen" = "#E69F00", "Oropharyngeal" = "gray60")

dominant_obs_grouped <- per_sample_ra_lab %>%
  filter(ra > 50) %>%
  mutate(group = factor(ifelse(group == "pathogen", "Pathogen", "Oropharyngeal"),
                        levels = c("Pathogen", "Oropharyngeal")))

n_per_group <- dominant_obs_grouped %>%
  count(group) %>%
  mutate(group_long = recode(as.character(group),
                             Pathogen      = "Pathogen-Dominated",
                             Oropharyngeal = "Oropharyngeal-Dominated"),
         label = paste0(group_long, "<br/>(n=", scales::comma(n), ")"))
group_label_map <- setNames(n_per_group$label, as.character(n_per_group$group))

q2_p_value <- q2_wilcox$p.value
q2_p_label <- if (q2_p_value < 2.2e-16) {
  "italic(P) < '2.2e-16'"
} else {
  sprintf("italic(P) == '%.1g'", q2_p_value)
}

panel_S1b_outliers <- dominant_obs_grouped %>%
  group_by(group) %>%
  mutate(.q1 = quantile(ra, 0.25, na.rm = TRUE),
         .q3 = quantile(ra, 0.75, na.rm = TRUE),
         .iqr = .q3 - .q1) %>%
  filter(ra < .q1 - 1.5 * .iqr | ra > .q3 + 1.5 * .iqr) %>%
  ungroup()

panel_S1b <- ggplot(dominant_obs_grouped, aes(x = group, y = ra, fill = group)) +
  geom_boxplot(alpha = 0.7, width = 0.5, notch = TRUE, outlier.shape = NA) +
  geom_point(data = panel_S1b_outliers, aes(fill = group),
             shape = 21, color = "black", stroke = 0.2,
             size = 1, alpha = 0.85) +
  scale_fill_manual(values = group_cols, guide = "none") +
  scale_x_discrete(labels = group_label_map) +
  scale_y_continuous(limits = c(45, 110), breaks = seq(50, 100, by = 10),
                     expand = expansion(mult = c(0, 0))) +
  annotate("segment", x = 1, xend = 2, y = 104, yend = 104,
           color = "black", linewidth = 0.4) +
  annotate("segment", x = 1, xend = 1, y = 102, yend = 104,
           color = "black", linewidth = 0.4) +
  annotate("segment", x = 2, xend = 2, y = 102, yend = 104,
           color = "black", linewidth = 0.4) +
  annotate("text", x = 1.5, y = 107, label = q2_p_label,
           parse = TRUE, size = 2.8, color = "black") +
  labs(x = NULL, y = "Relative Abundance (%)") +
  fig1_theme +
  theme(axis.text.x = ggtext::element_markdown(angle = 45, hjust = 1, size = 10))
```

### Figure S1

```{r fig_S1, include=FALSE}
figure_S1 <- panel_S1a + panel_S1b +
  patchwork::plot_layout(widths = c(1.6, 1)) +
  patchwork::plot_annotation(
    tag_levels = list(c("A", "B")),
    theme = theme(plot.tag = element_text(size = 12, face = "bold"))
  )

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S1.pdf", figure_S1, units = "in",
       width = 9, height = 5, device = cairo_pdf)
```

### Figure S2

```{r fig_S2, include=FALSE}
if (!"Age_bin" %in% colnames(nmds_scores)) {
  nmds_scores$Age_bin <- cut(nmds_scores$Age,
    breaks = c(-Inf, 2, 12, 18, 40, Inf), right = FALSE,
    labels = c("Infant", "Child", "Adolescent", "Adult", "Senior"))
}
if (!"FEV1_bin" %in% colnames(nmds_scores)) {
  nmds_scores$FEV1_bin <- cut(nmds_scores$FEV1,
    breaks = c(-Inf, 30, 50, 80, Inf), right = FALSE,
    labels = c("Very Severe", "Severe", "Moderate", "Mild"))
}

s1_base <- theme_classic() +
  theme(
    axis.text        = element_text(color = "black", size = 8),
    axis.title       = element_text(color = "black", size = 10),
    axis.title.x     = element_text(margin = margin(t = 0)),
    axis.title.y     = element_text(margin = margin(r = 3)),
    axis.ticks       = element_line(color = "black", linewidth = 0.25),
    axis.line        = element_blank(),
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.25),
    plot.background  = element_blank(),
    plot.margin      = margin(2, 5, 2, 2),
    plot.tag         = element_text(size = 12, face = "bold"),
    legend.position  = "bottom",
    legend.title     = element_text(size = 7, face = "bold"),
    legend.text      = element_text(size = 6),
    legend.key.size  = unit(0.25, "cm"),
    legend.margin    = margin(0, 0, 0, 0),
    legend.box.margin = margin(-3, 0, 0, 0)
  )

age_levels_s1   <- c("Infant", "Child", "Adolescent", "Adult", "Senior")
age_range_s1    <- c("Infant"     = "0\u20131",
                     "Child"      = "2\u201311",
                     "Adolescent" = "12\u201317",
                     "Adult"      = "18\u201339",
                     "Senior"     = "40+")
age_cols_s1     <- c("Infant" = "#F0E442", "Child" = "#E69F00",
                     "Adolescent" = "#CC79A7", "Adult" = "#4575B4",
                     "Senior" = "#56B4E9")
fev1_levels_s1  <- c("Mild", "Moderate", "Severe", "Very Severe")
fev1_range_s1   <- c("Mild" = "\u226580%", "Moderate" = "50\u201379%",
                     "Severe" = "30\u201349%", "Very Severe" = "<30%")
fev1_cols_s1    <- c("Mild" = "#4575B4", "Moderate" = "#ABD9E9",
                     "Severe" = "#FDAE61", "Very Severe" = "#D73027")

stress_lbl <- paste0("Stress == ", signif(nmds_stress, 2))

legend_with_n <- function(values, label_map) {
  counts <- table(factor(values, levels = names(label_map)))
  setNames(paste0(label_map[names(counts)], " (N=", counts, ")"),
           names(counts))
}

s1a_df <- nmds_scores %>%
  filter(!is.na(Age_bin)) %>%
  mutate(Age_bin = factor(Age_bin, levels = age_levels_s1))
age_labels_s1  <- legend_with_n(s1a_df$Age_bin, age_range_s1)

panel_S2a <- ggplot(s1a_df, aes(MDS1, MDS2, color = Age_bin)) +
  geom_point(size = 0.5, alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.95, linewidth = 0.4) +
  scale_color_manual(values = age_cols_s1, labels = age_labels_s1,
                      name = "Age (years)", drop = FALSE) +
  annotate("text", x = Inf, y = -Inf, label = stress_lbl, parse = TRUE,
           hjust = 1.05, vjust = -0.5, size = 2) +
  labs(x = "NMDS1", y = "NMDS2", tag = "A") +
  s1_base +
  theme(aspect.ratio = 1) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE,
                               override.aes = list(size = 2, alpha = 1)))

s1b_df <- nmds_scores %>%
  filter(!is.na(FEV1_bin)) %>%
  mutate(FEV1_bin = factor(FEV1_bin, levels = fev1_levels_s1))
fev1_labels_s1 <- legend_with_n(s1b_df$FEV1_bin, fev1_range_s1)

panel_S2b <- ggplot(s1b_df, aes(MDS1, MDS2, color = FEV1_bin)) +
  geom_point(size = 0.5, alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.95, linewidth = 0.4) +
  scale_color_manual(values = fev1_cols_s1, labels = fev1_labels_s1,
                      name = expression(FEV[1]~pp), drop = FALSE) +
  annotate("text", x = Inf, y = -Inf, label = stress_lbl, parse = TRUE,
           hjust = 1.05, vjust = -0.5, size = 2) +
  labs(x = "NMDS1", y = "NMDS2", tag = "B") +
  s1_base +
  theme(aspect.ratio = 1) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 2, alpha = 1)))

s1c_df <- nmds_scores %>%
  filter(!is.na(Sex) & Sex != "")
sex_labels_s1 <- setNames(
  paste0(names(table(s1c_df$Sex)), " (N=", as.integer(table(s1c_df$Sex)), ")"),
  names(table(s1c_df$Sex))
)

panel_S2c <- ggplot(s1c_df, aes(MDS1, MDS2, color = Sex)) +
  geom_point(size = 0.5, alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.95, linewidth = 0.4) +
  scale_color_brewer(palette = "Set1", labels = sex_labels_s1, name = "Sex") +
  annotate("text", x = Inf, y = -Inf, label = stress_lbl, parse = TRUE,
           hjust = 1.05, vjust = -0.5, size = 2) +
  labs(x = "NMDS1", y = "NMDS2", tag = "C") +
  s1_base +
  theme(aspect.ratio = 1) +
  guides(color = guide_legend(nrow = 1, override.aes = list(size = 2, alpha = 1)))

figure_S2 <- panel_S2a | panel_S2b | panel_S2c

ggsave("figures/figure_S2.pdf", figure_S2, width = 9, height = 6, device = cairo_pdf)
```

### Figure S3

```{r fig_S3_genus_age, include=FALSE}
GENUS_THRESH <- 1     

ga_build <- function(require_fev1) {
  M <- genus_matrix_filt
  M <- M[rowSums(M) > 0, , drop = FALSE]
  mt <- cf_microbiome_metadata[match(rownames(M), cf_microbiome_metadata$Run), ]
  mt$FEV1 <- suppressWarnings(as.numeric(mt$FEV1))
  mt$Age  <- suppressWarnings(as.numeric(mt$Age))
  ok <- !is.na(mt$Age) & !is.na(mt$Sample_Type) & mt$Sample_Type != ""
  if (require_fev1) ok <- ok & !is.na(mt$FEV1)
  M <- M[ok, , drop = FALSE]; mt <- mt[ok, ]
  ind <- run_indiv[rownames(M)]
  keep <- !is.na(ind)
  M <- M[keep, , drop = FALSE]; mt <- mt[keep, ]; ind <- ind[keep]
  pool <- rowsum(M, group = ind)
  modal <- function(x) names(sort(table(x), decreasing = TRUE))[1]
  meta <- data.frame(
    indiv = rownames(pool),
    Age   = tapply(mt$Age,         ind, mean)[rownames(pool)],
    FEV1  = tapply(mt$FEV1,        ind, mean)[rownames(pool)],
    BP    = tapply(mt$BioProject,  ind, modal)[rownames(pool)],
    ST    = tapply(mt$Sample_Type, ind, modal)[rownames(pool)],
    stringsAsFactors = FALSE)
  list(ra = sweep(pool, 1, rowSums(pool), "/") * 100, meta = meta)
}
ga_age <- ga_build(FALSE)
ga_lf  <- ga_build(TRUE)

ga_age$meta$agr <- factor(ifelse(ga_age$meta$Age >= 18, "18+", "0–17"),
                          levels = c("0–17", "18+"))
ga_lf$meta$fev10 <- -ga_lf$meta$FEV1 / 10   # per 10-point DECREASE

ga_fit <- function(obj, term, adj_age) {
  do.call(rbind, lapply(colnames(obj$ra), function(g) {
    d <- data.frame(pres = as.integer(obj$ra[, g] >= GENUS_THRESH),
                    x = obj$meta[[term]], BP = factor(obj$meta$BP),
                    ST = factor(obj$meta$ST), Age = obj$meta$Age)
    d <- d[complete.cases(d), ]
    f <- if (adj_age) pres ~ x + Age + BP + ST else pres ~ x + BP + ST
    s <- tryCatch(summary(glm(f, data = d, family = binomial))$coefficients,
                  error = function(e) NULL)
    rn <- if (is.null(s)) NA else grep("^x", rownames(s), value = TRUE)[1]
    if (is.null(s) || is.na(rn)) return(NULL)
    data.frame(genus = g, OR = exp(s[rn, 1]),
               lo = exp(s[rn, 1] - 1.96 * s[rn, 2]),
               hi = exp(s[rn, 1] + 1.96 * s[rn, 2]),
               p = s[rn, 4], stringsAsFactors = FALSE)
  }))
}
ga_A <- ga_fit(ga_age, "agr",   FALSE)   
ga_B <- ga_fit(ga_lf,  "fev10", TRUE)    
ga_A$q <- p.adjust(ga_A$p, "BH")
ga_B$q <- p.adjust(ga_B$p, "BH")

dir.create("tables", showWarnings = FALSE)

ga_gen  <- colnames(ga_age$ra)
ga_ord  <- c(fig1_pathogen_order[fig1_pathogen_order %in% ga_gen],
             sort(ga_gen[!ga_gen %in% cf_pathogens]))
ga_gold <- "#E69F00"
ga_labs <- setNames(ifelse(rev(ga_ord) %in% cf_pathogens,
  paste0("<span style='color:", ga_gold, "'><i>", rev(ga_ord), "</i></span>"),
  paste0("<i>", rev(ga_ord), "</i>")), rev(ga_ord))

ga_prep <- function(d, mult) {
  d$sig  <- d$q < 0.05
  d$path <- d$genus %in% cf_pathogens
  qv <- ifelse(d$q < 0.001, formatC(d$q, format = "e", digits = 1),
        ifelse(d$q < 0.01,  sprintf("%.3f", d$q), sprintf("%.2f", d$q)))
  d$txt <- paste0("*Q* = ", qv)
  d$qx  <- ifelse(d$OR < 1, d$lo / mult, d$hi * mult)
  d$qh  <- ifelse(d$OR < 1, 1, 0)
  d$genus <- factor(d$genus, levels = rev(ga_ord))
  d
}
ga_A <- ga_prep(ga_A, 1.07); ga_B <- ga_prep(ga_B, 1.015)

ga_panel <- function(d, brks, labl, lims, xlab, ylabs) {
  p <- ggplot(d, aes(OR, genus, color = path)) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.5) +
    geom_errorbar(aes(xmin = lo, xmax = hi), width = 0, linewidth = 0.6,
                  orientation = "y") +
    geom_point(aes(shape = sig), size = 2.2, fill = "white", stroke = 0.9) +
    ggtext::geom_richtext(aes(x = qx, label = txt, hjust = qh), size = 2.8,
                          color = "black", fill = NA, label.color = NA,
                          label.padding = unit(0, "pt")) +
    scale_color_manual(values = c(`TRUE` = ga_gold, `FALSE` = "black"), guide = "none") +
    scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 21), guide = "none") +
    scale_x_log10(breaks = brks, labels = labl, limits = lims,
                  expand = expansion(mult = c(0.01, 0.01))) +
    labs(x = xlab, y = NULL) +
    gg_theme +
    theme(aspect.ratio = 1,
          axis.text.x  = element_text(size = 9.5, color = "black"),
          axis.title.x = ggtext::element_markdown(size = 11, color = "black"),
          axis.line    = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
          plot.tag     = element_text(size = 13, face = "bold"))
  if (ylabs) p + scale_y_discrete(labels = ga_labs) +
                 theme(axis.text.y = ggtext::element_markdown(size = 9.5))
  else       p + scale_y_discrete(labels = rep("", length(ga_ord))) +
                 theme(axis.text.y  = element_blank(),
                       axis.ticks.y = element_line(color = "black", linewidth = 0.25))
}
ga_arrA <- paste0("<span style='font-size:9pt'>&#8592; Enriched in Children",
  "<span style='color:white'>", strrep("n", 5), "</span>",
  "Enriched in Adults &#8594;</span>")
ga_arrB <- paste0("<span style='font-size:9pt'>&#8592; Better Lung Function",
  "<span style='color:white'>", strrep("n", 5), "</span>",
  "Worse Lung Function &#8594;</span>")

ga_pA <- ga_panel(ga_A, c(0.25, 0.5, 1, 2, 4, 8), c("0.25", "0.5", "1", "2", "4", "8"),
  c(0.11, 17),
  paste0("Odds Ratio for Colonization<br>(Adults vs. Children)<br>", ga_arrA),
  TRUE) + labs(tag = "A")
ga_pB <- ga_panel(ga_B, c(0.75, 1, 1.25, 1.5), c("0.75", "1", "1.25", "1.5"),
  c(0.63, 1.80),
  paste0("Odds Ratio for Colonization<br>(per 10-Point Decrease in FEV<sub>1</sub> pp)<br>",
         ga_arrB),
  FALSE) + labs(tag = "B")

figure_S3 <- ga_pA + ga_pB + patchwork::plot_layout(widths = c(1, 1))

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S3.pdf", figure_S3,
       width = 12.5, height = 6.4, device = cairo_pdf)

sprintf("A: %d individuals, %d of %d genera significant",
        nrow(ga_age$ra), sum(ga_A$sig), nrow(ga_A))
sprintf("B: %d individuals, %d of %d genera significant",
        nrow(ga_lf$ra),  sum(ga_B$sig), nrow(ga_B))
```

### Figure S4 

```{r fig_S4_analysis, include=FALSE}
MIN_REGION_N <- 100

region_vec <- cf_microbiome_metadata$Region[
  match(rownames(genus_matrix_filt), cf_microbiome_metadata$Run)]

overall_pairs <- pairwise_results[, c("genus1", "genus2", "bx", "pval", "significant")]
names(overall_pairs)[names(overall_pairs) == "significant"] <- "sig"
overall_pairs$r <- mapply(function(a, b) cor(log_abund[, a], log_abund[, b]),
                          overall_pairs$genus1, overall_pairs$genus2)
overall_pairs$direction <- ifelse(overall_pairs$bx > 0, "Positive", "Negative")

region_levels <- sort(unique(region_vec))

region_pair_list   <- list()
region_concordance <- list()
for (rg in region_levels) {
  idx <- which(region_vec == rg)
  rr  <- fit_region_pairs(log10(genus_matrix_filt[idx, , drop = FALSE] + 1))
  rr$sig <- rr$pval < bonf_thresh

  m <- merge(overall_pairs[, c("genus1", "genus2", "bx", "r", "sig", "direction")],
             rr[, c("genus1", "genus2", "bx", "r", "sig")],
             by = c("genus1", "genus2"), suffixes = c("_ov", "_rg"))
  m$region <- rg
  m$n      <- length(idx)
  region_pair_list[[rg]] <- m

  est   <- !is.na(m$bx_rg)                       # pair estimable in this region
  ovsig <- m$sig_ov & est                        # associations "predicted" by the pooled model
  region_concordance[[rg]] <- data.frame(
    region           = rg,
    n                = length(idx),
    n_estimable      = sum(est),
    n_sig_region     = sum(m$sig_rg & est, na.rm = TRUE),
    n_overall_sig    = sum(ovsig),
    sign_concordance = 100 * mean(sign(m$bx_rg[ovsig]) == sign(m$bx_ov[ovsig])),
    spearman_bx      = cor(m$bx_ov[est], m$bx_rg[est], method = "spearman"),
    ccc_bx           = ccc(m$bx_ov[est], m$bx_rg[est]),
    pearson_std_r    = cor(m$r_ov[est],  m$r_rg[est],  method = "pearson"),
    stringsAsFactors = FALSE)
}
region_pair_df     <- do.call(rbind, region_pair_list)
region_concordance <- do.call(rbind, region_concordance)
region_concordance$underpowered <- region_concordance$n < MIN_REGION_N

dir.create("tables", showWarnings = FALSE)
print(region_concordance, row.names = FALSE)
```

```{r fig_S4, include=FALSE}
region_lab <- setNames(
  sprintf("%s (N=%s)", region_concordance$region,
          scales::comma(region_concordance$n)),
  region_concordance$region)

assoc_cols <- c("Positive" = "#4575B4", "Negative" = "#D73027",
                "NS" = "gray75")

AXIS_LIM <- range(c(region_pair_df$bx_ov, region_pair_df$bx_rg), na.rm = TRUE)
AXIS_LIM <- AXIS_LIM + c(-1, 1) * 0.06 * diff(AXIS_LIM)

pair_plot_df <- region_pair_df
pair_plot_df$region_f <- factor(region_lab[pair_plot_df$region],
                                levels = region_lab[region_levels])
pair_plot_df$pt_class <- factor(
  ifelse(!pair_plot_df$sig_ov, "NS", pair_plot_df$direction),
  levels = c("Positive", "Negative", "NS"))
pair_plot_df <- pair_plot_df[order(pair_plot_df$pt_class, decreasing = TRUE), ]

facet_ann <- region_concordance
facet_ann$region_f <- factor(region_lab[facet_ann$region],
                             levels = region_lab[region_levels])
facet_ann$txt <- sprintf("Signs %.0f%%  ·  CCC=%.2f",
                         facet_ann$sign_concordance, facet_ann$ccc_bx)

panel_S4A <- ggplot(pair_plot_df, aes(x = bx_ov, y = bx_rg)) +
  geom_hline(yintercept = 0, color = "gray85", linewidth = 0.25) +
  geom_vline(xintercept = 0, color = "gray85", linewidth = 0.25) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "gray50", linewidth = 0.4) +
  geom_point(aes(fill = pt_class), shape = 21, color = "black",
             stroke = 0.15, size = 1.5, alpha = 0.85) +
  geom_text(data = facet_ann, aes(x = AXIS_LIM[1] + 0.03, y = AXIS_LIM[2] - 0.03,
            label = txt), hjust = 0, vjust = 1, size = 2.7, color = "black",
            inherit.aes = FALSE) +
  facet_wrap(~ region_f, ncol = 4) +
  scale_fill_manual(values = assoc_cols, name = NULL) +
  scale_x_continuous(breaks = c(-2, -1, 0, 1)) +
  scale_y_continuous(breaks = c(-2, -1, 0, 1)) +
  coord_cartesian(xlim = AXIS_LIM, ylim = AXIS_LIM) +
  labs(x = "Overall Covariation Coefficient",
       y = "Per-Region Covariation Coefficient",
       tag = "A") +
  gg_theme +
  theme(axis.text.x  = element_text(size = 10, color = "black"),
        axis.text.y  = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        strip.text   = element_text(size = 11, color = "black"),
        strip.background = element_blank(),
        legend.position = "bottom",
        legend.text  = element_text(size = 10),
        plot.tag     = element_text(size = 12, face = "bold")) +
  guides(fill = guide_legend(override.aes = list(size = 3, alpha = 1)))

metric_levels <- c("Signs Matched", "Spearman Rank", "CCC")
metric_df <- rbind(
  data.frame(region = region_concordance$region, metric = metric_levels[1],
             value = region_concordance$sign_concordance / 100),
  data.frame(region = region_concordance$region, metric = metric_levels[2],
             value = region_concordance$spearman_bx),
  data.frame(region = region_concordance$region, metric = metric_levels[3],
             value = region_concordance$ccc_bx))
metric_df$metric <- factor(metric_df$metric, levels = metric_levels)

metric_df$region_f <- factor(metric_df$region, levels = rev(region_levels))

metric_cols  <- setNames(c("#4575B4", "#E69F00", "#333333"), metric_levels)
metric_shape <- setNames(c(16, 17, 15), metric_levels)

panel_S4B <- ggplot(metric_df, aes(x = value, y = region_f,
                                   color = metric, shape = metric)) +
  geom_point(size = 2.6, position = position_dodge(width = 0.6)) +
  scale_color_manual(values = metric_cols, name = NULL) +
  scale_shape_manual(values = metric_shape, name = NULL) +
  scale_x_continuous(breaks = seq(0.4, 1, 0.2)) +
  coord_cartesian(xlim = c(min(metric_df$value) - 0.03, 1)) +
  labs(x = "Concordance With Pooled Analysis", y = NULL, tag = "B") +
  gg_theme +
  theme(axis.text.x        = element_text(size = 10, color = "black"),
        axis.text.y        = element_text(size = 10, color = "black"),
        axis.title.x       = element_text(size = 12, color = "black"),
        legend.position    = "bottom",
        legend.text        = element_text(size = 10),
        plot.tag           = element_text(size = 12, face = "bold"),
        panel.grid.major.x = element_line(color = "gray92", linewidth = 0.25))

figure_S4 <- panel_S4A / panel_S4B +
  patchwork::plot_layout(heights = c(1.5, 1))

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S4.pdf", figure_S4, width = 9, height = 8.5,
       device = cairo_pdf)
```

### Figure S5

```{r fig_S5, include=FALSE}
present_pathogens <- sort(intersect(fig1_pathogen_order, unique(bp_df$dominant_pathogen)))
present_pathogens <- setdiff(present_pathogens, "Haemophilus")
dom_path_levels   <- c(present_pathogens, "Other")
dom_path_cols     <- c(pathogen_palette[present_pathogens], "Other" = "gray60")

bp_plot_df_s2 <- bp_adj_df %>%
  mutate(
    dominant_pathogen = if_else(dominant_pathogen == "Haemophilus",
                                "Other", dominant_pathogen),
    dominant_pathogen = factor(dominant_pathogen, levels = dom_path_levels)
  )

dom_path_counts <- bp_plot_df_s2 %>% count(dominant_pathogen)
dom_path_labels <- setNames(
  vapply(dom_path_levels, function(g) {
    n <- dom_path_counts$n[dom_path_counts$dominant_pathogen == g]
    if (length(n) == 0) n <- 0
    if (g == "Other") {
      paste0("Other (N=", scales::comma(n), ")")
    } else {
      paste0("<i>", g, "</i> (N=", scales::comma(n), ")")
    }
  }, character(1)),
  dom_path_levels
)

panel_S5_base <- ggplot(bp_plot_df_s2, aes(x = bp_dom, y = FEV1_adj)) +
  geom_line(data = boot_lines_bp, aes(x = x, y = y, group = boot),
            color = "lightpink", alpha = 0.15) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_jitter(aes(fill = dominant_pathogen),
              shape = 21, color = "black", stroke = 0.2,
              width = 0.005, alpha = 0.85, size = 1.5) +
  geom_smooth(method = "lm", color = "#D73027", se = FALSE, linewidth = 1.1) +
  scale_fill_manual(values = dom_path_cols, labels = dom_path_labels,
                    name = "Most Abundant Genus", drop = FALSE) +
  scale_x_continuous(breaks = seq(0.2, 1.0, 0.2)) +
  scale_y_continuous() +
  labs(x = "Dominance (BP)",
       y = expression(Lung~Function~"("*FEV[1]~pp[Adj]*")")) +
  fig1_theme +
  theme(legend.position   = "bottom",
        legend.text       = ggtext::element_markdown(size = 9),
        legend.title      = element_text(size = 10, face = "bold"),
        legend.key.height = unit(0.35, "cm"),
        legend.margin     = margin(t = 4, b = 0),
        aspect.ratio      = NULL) +
  guides(fill = guide_legend(nrow = 4, ncol = 2, byrow = TRUE,
                             override.aes = list(alpha = 1, size = 1.5, shape = 21)))

panel_S5_fixed <- egg::set_panel_size(panel_S5_base,
                                       width  = fig2_panel_size,
                                       height = fig2_panel_size)

figure_S5 <- panel_S5_fixed

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S5.pdf", figure_S5,
       width = 7, height = 6, device = cairo_pdf)
```

### Figure S6

```{r fig_S6, include=FALSE}
hid_bold_S4       <- c("HID7", "HID8")
isolate_hilite_S4 <- c("NB0017", "NB0078", "NB1046", "NB0219", "RSD0018", "NB1386")

hid_order_S4 <- pseudomonas_variable_isolates %>%
  arrange(desc(z_spread)) %>%
  pull(hid)

s4_df <- total_inhibition_per_isolate %>%
  filter(grepl("Pseudomonas", genus, ignore.case = TRUE),
         hid %in% hid_order_S4) %>%
  mutate(
    hid       = factor(hid, levels = hid_order_S4),
    highlight = isolate %in% isolate_hilite_S4
  )

pad_hid_S4 <- function(h) {
  num <- sub("^HID", "", h)
  sprintf("HID%02d", as.integer(num))
}

hid_axis_labels_S4 <- setNames(
  vapply(hid_order_S4, function(h) {
    lab <- pad_hid_S4(h)
    if (h %in% hid_bold_S4) paste0("**", lab, "**") else lab
  }, character(1)),
  hid_order_S4
)

panel_S6 <- ggplot(s4_df, aes(x = hid, y = z_within_genus)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_boxplot(outlier.shape = NA, fill = "gray92",
               color = "black", linewidth = 0.3, width = 0.6) +
  geom_jitter(aes(color = highlight),
                            cex = 2, size = 1.6, alpha = 0.85) +
  scale_color_manual(values = c(`FALSE` = "gray40", `TRUE` = "#E69F00"),
                      guide = "none") +
  scale_x_discrete(labels = hid_axis_labels_S4) +
  labs(x = "Donor ID", y = "Within-Genus Inhibition Z-Score") +
  fig1_theme +
  theme(axis.text.x = ggtext::element_markdown(angle = 45, hjust = 1, size = 10))

figure_S6 <- panel_S6

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S6.pdf", figure_S6,
       width = 9, height = 4.5, device = cairo_pdf)
```

### Figure S7B

```{r fig_S7B, include=FALSE}
lcms_iso_sub <- t(apply(lcms_sub, 1, function(v) tapply(v, lcms_iso, mean)))
lcms_iprof   <- t(lcms_iso_sub); lcms_iprof <- lcms_iprof / rowSums(lcms_iprof)
dir.create("r_objects", showWarnings = FALSE)
if (file.exists("r_objects/lcms_permanova.rds")) {
  lcms_perm <- readRDS("r_objects/lcms_permanova.rds")
} else {
  set.seed(8675309)   # PERMANOVA permutation reproducibility
  lcms_perm <- adonis2(vegdist(lcms_iprof, "bray") ~ host + inhibition,
                       data = lcms_iso_meta, by = "margin", permutations = lcms_n_perm)
  saveRDS(lcms_perm, "r_objects/lcms_permanova.rds")
}
lcms_msres <- lcms_perm["Residual", "SumOfSqs"] / lcms_perm["Residual", "Df"]
lcms_sstot <- lcms_perm["Total", "SumOfSqs"]
lcms_omega2 <- function(tm) (lcms_perm[tm, "SumOfSqs"] - lcms_perm[tm, "Df"] * lcms_msres) /
  (lcms_sstot + lcms_msres)
lcms_eff <- data.frame(
  term   = c("host (donor)", "inhibition"),
  Df     = c(lcms_perm["host", "Df"], lcms_perm["inhibition", "Df"]),
  R2     = c(lcms_perm["host", "R2"], lcms_perm["inhibition", "R2"]),
  omega2 = c(lcms_omega2("host"), lcms_omega2("inhibition")),
  F      = c(lcms_perm["host", "F"], lcms_perm["inhibition", "F"]),
  p      = c(lcms_perm["host", "Pr(>F)"], lcms_perm["inhibition", "Pr(>F)"]))

# PCoA on per-injection profiles (Bray-Curtis, total-sum-scaled)
lcms_sprof <- t(lcms_sub); lcms_sprof <- lcms_sprof / rowSums(lcms_sprof)
lcms_pcoa  <- cmdscale(vegdist(lcms_sprof, "bray"), k = 2, eig = TRUE)
lcms_pvar  <- round(100 * lcms_pcoa$eig[1:2] / sum(lcms_pcoa$eig[lcms_pcoa$eig > 0]), 1)
lcms_mdsdf <- data.frame(lcms_samp_meta, PCoA1 = lcms_pcoa$points[, 1], PCoA2 = lcms_pcoa$points[, 2])

figure_S7B <- ggplot(lcms_mdsdf, aes(PCoA1, PCoA2)) +
  stat_ellipse(aes(color = host, group = host), type = "t", linewidth = 0.4, show.legend = FALSE) +
  stat_ellipse(data = subset(lcms_mdsdf, inhibition == "Yes"), color = lcms_inh_pal[["Yes"]],
               type = "t", linetype = "22", linewidth = 1) +
  stat_ellipse(data = subset(lcms_mdsdf, inhibition == "No"), color = lcms_inh_pal[["No"]],
               type = "t", linetype = "22", linewidth = 1) +
  geom_point(aes(color = host, shape = inhibition), size = 3, stroke = 0.6) +
  scale_color_manual(values = lcms_host_pal, name = "HID") +
  scale_shape_manual(values = c(Yes = 16, No = 17), name = "Inhibition\nZ-Score ≥ 0") +
  labs(x = sprintf("PCoA1 (%.1f%%)", lcms_pvar[1]), y = sprintf("PCoA2 (%.1f%%)", lcms_pvar[2])) +
  coord_fixed(ratio = 1) + fig1_theme
lcms_xy  <- do.call(rbind, lapply(ggplot_build(figure_S7B)$data, function(d) d[c("x", "y")]))
lcms_lim <- range(lcms_xy, na.rm = TRUE); lcms_lim <- lcms_lim + c(-0.03, 0.03) * diff(lcms_lim)
figure_S7B <- figure_S7B + coord_fixed(ratio = 1, xlim = lcms_lim, ylim = lcms_lim)

dir.create("figures", showWarnings = FALSE)
ggsave("figures/figure_S7B.pdf", figure_S7B, width = 6.5, height = 6.5, device = cairo_pdf)
```
